---
title: "Labor Market Analysis"
author: "Marjorie Blanco, Joe Thomson, Haodi Tu"
subtitle: Exploratory Data Analysis
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r, echo=FALSE}
# Course: ECON 5330
# Title: EDA Assignment
# Date: Jan 22, 2019
# Author: Haodi Tu, Majorie Blanco, Joseph Thomson
```

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(readr)
library(ggplot2)
library(mosaic)
library(dplyr)
library(kableExtra)
library(qwraps2)
```

## Research Questions and Geographic area 

## Data

<!-- https://cran.r-project.org/web/packages/qwraps2/vignettes/summary-statistics.html -->

```{r}
ss16ppr <- read_csv("csv_ppr/ss16ppr.csv")
#kable(ss16ppr)
```

We used data from the 2016 ACS for Puerto Rico to examine earnings gaps between individuals with different education levels. Our research questions are:
1) How do earnings vary by education level?
2) How does the premium for education vary by gender?
The 2016 ACS is a nationally representative sample of `r nrow(ss16ppr)`. The household survey includes questions pertaining to each household member’s demographic characteristics and labor market activity. 

We restrict our sample to these three racial groups: White, Black and Other. In addition, given our goal of examining earnings differences by ability to speak English and the reporting of earnings in the ACS on an annual basis (wages, salary, commissions, bonuses, tips, and self-employment income during the past 12 months), we restrict our sample to full-time year-round (FTYR) workers. We define FTYR workers as individuals who report positive earnings over the past year, who worked at least 40 of the past 52 weeks, and who worked at least 35 hours per week in a usual work week over this period. 

<!-- Our sample consists of 1,443 deaf or hard-of-hearing men, 172,048 men without a hearing disability, 728 deaf or hard-of-hearing women, and 134,877 women without a hearing disability. -->


```{r}
# add column labels
attr(ss16ppr$RT, "label") <- "Record Type"
attr(ss16ppr$RAC1P, "label") <- "Recoded detailed race code"
attr(ss16ppr$ENG, "label") <- "Ability to speak English"
attr(ss16ppr$SCHL, "label") <- "Educational attainment"
```

```{r}
#select columns of interest
ss16ppr <- ss16ppr %>% select(WAGP, SEX, AGEP, RAC1P, RACWHT, RACBLK, HISP, MAR, SCHL, ENG, COW, WKW, WKHP)
```

## Sample selection criteria

```{r}
ss16ppr$WKHP <- as.numeric(ss16ppr$WKHP)
ss16ppr$WAGP <- as.numeric(ss16ppr$WAGP)
ss16ppr$AGEP <- as.numeric(ss16ppr$AGEP)


#Filters
# AGE: Only between 18 and 64 (inclusive) years of age
ss16ppr <- ss16ppr %>% filter(AGEP >= 18, AGEP <= 64)

# SCHL: High school and above
#       Combine: 16/17 for High school and 18/19 for Some college
ss16ppr <- ss16ppr %>% filter(SCHL >= 16) %>% mutate(SCHL = ifelse(SCHL == 17, 16, SCHL)) %>% mutate(SCHL = ifelse(SCHL == 19, 18, SCHL))

# WKW: Only between 50 and 40 hours
# 1 .50 to 52 weeks worked during past 12 months
# 2 .48 to 49 weeks worked during past 12 months
# 3 .40 to 47 weeks worked during past 12 months
ss16ppr <- ss16ppr %>% filter(between(WKW, 1,3))

# Remove NAs and zero in WAGP
ss16ppr <- ss16ppr %>% filter(!is.na(WAGP))
ss16ppr <- ss16ppr %>% filter(WAGP != 0) 
# Remove outlier
ss16ppr <-ss16ppr %>% filter(WAGP != max(ss16ppr$WAGP))

# Restirct class of worker to exclude those working without pay and unemployed
# Class of worker
# b N/A (less than 16 years old/NILF who last worked more than 5 years ago or never worked)
# 1 Employee of a private for-profit company or business, or of an individual, for wages,
#    salary, or commissions
# 2 Employee of a private not-for-profit, tax-exempt, or charitable organization
# 3 Local government employee (city, county, etc.)
# 4 State government employee
# 5 Federal government employee
# 6 Self-employed in own not incorporated business, professional practice, or farm
# 7 Self-employed in own incorporated business, professional practice or farm
# 8 Working without pay in family business or farm
# 9 Unemployed and last worked 5 years ago or earlier or never worked
ss16ppr <- ss16ppr %>% filter(between(COW, 1, 7))


# WKHP: worked at least 35 hr per week in a usual work week over this period. 
ss16ppr <- ss16ppr %>% filter(WKHP >= 35)
```


```{r}
# Decode variables of interest
ss16ppr  <- ss16ppr %>% mutate(RACHISP = ifelse(HISP == "01", "Yes", "No"))
ss16ppr  <- ss16ppr %>% mutate(RACWHT = ifelse(RACWHT == 1, "Yes", "No"))
ss16ppr  <- ss16ppr %>% mutate(RACBLK = ifelse(RACBLK == 1, "Yes", "No"))
ss16ppr <- ss16ppr %>% mutate(MAR1 = ifelse(MAR == 1, "Yes", "No"))

#code to decode MAR
ss16ppr <- mutate(ss16ppr, MAR = derivedFactor(
  "Married" = MAR == 1,
  "Widowed" = MAR == 2,
  "Divorced" = MAR == 3,
  "Separated" = MAR == 4,
  "Never married" = MAR == 5,
  method ="first",
  .default = "NA"
))

#code to decode ENG
ss16ppr <- mutate(ss16ppr, ENG = derivedFactor(
  "NA" = (ENG == "1" | is.na(ENG)),
  "Very well" = ENG == "2",
  "Well" = ENG == "3",
  "Not well" = ENG == "4",
  "Not at all" = ENG == "5",
  method ="first",
  .default = "NA"
))

#Note: 16 .Regular high school diploma

#code to decode SCHL
ss16ppr <- mutate(ss16ppr, SCHL = derivedFactor(
  "High school diploma" = SCHL == "16",
  "Some college" = SCHL == "18",
  "Associate's degree" = SCHL == "20",
  "Bachelor's degree" = SCHL == "21",
  "Master's degree" = SCHL == "22",
  "Professional degree" = SCHL == "23",
  "Doctorate degree" = SCHL == "24",
  method ="first",
  .default = "NA"
))
```
## EDA Insights: 

For our exploratory analysis we looked at population breakdowns by education, age,  marital status, gender, race, earnings, and work hours. We applied filters on education (HS diploma or above), age (18-64), and work hours (>35/week). 

An earnings histogram identified a default maximum amount of earnings (189k) which we also filtered out of the data. The wage distribution is progressive above the median, but drops off sharply below the median, likely indicating the presence of a minimum wage. The correlation between age and earnings is very weak (.23). Likewise, earnings is very weakly correlated with hours worked among those who work more than 35 hours per week.

The age distribution of full time workers is skewed towards older adults, possibly indicating that younger workers have trouble finding full-time work, wait to enter the workforce, or are leaving the territory.

A scatterplot of age and earnings shows the highest earners work 40-60 hours per week. Also, a comparison of race and earnings shows little difference in median earnings based on race.


## Task 1: 

Examine the first 10 or 20 observations (rows of data) corresponding to variables of interest (columns) and compare the observed values to the data dictionary for person records.

```{r}
kable(head(ss16ppr, 20)) %>%
  kable_styling(bootstrap_options = "striped")
```

## Task 2:

Compute and examine descriptive statistics including the minimum, maximum, mean, and median for quantitative variables of interest

### Wage

```{r}
ss16ppr$WAGP <- as.numeric(ss16ppr$WAGP)
summary1 <-
  list("Wages or salary income past 12 months" =
       list("Minimum" = ~ min(.data$WAGP),
            "Maximum" = ~ max(.data$WAGP),
            "Median" = ~ median(.data$WAGP),
            "Mean" = ~ mean(.data$WAGP))
            
#            "Mean" = ~ mean(.data$WAGP),
#            "Mean (sd)" = ~ qwraps2::mean_sd(.data$WAGP))
       )
### Overall
whole <- summary_table(ss16ppr, summary1)
kable(whole) %>%
  kable_styling(bootstrap_options = "striped")
```

### Age

```{r}
summary2 <-
  list("Age" =
       list("Minimum" = ~ min(.data$AGEP),
            "Maximum" = ~ max(.data$AGEP),
            "Median" = ~ median(.data$AGEP),
             "Mean" = ~ mean(.data$WAGP))
            # "Mean" = ~ mean(.data$WAGP),
            # "Mean (sd)" = ~ qwraps2::mean_sd(.data$AGEP))
       )
### Overall
whole <- summary_table(ss16ppr, summary2)
kable(whole) %>%
  kable_styling(bootstrap_options = "striped")
```

### Hours worked 

```{r}
summary3 <-
  list("Usual hours worked per week past 12 months" =
       list("Minimum" = ~ min(.data$WKHP),
            "Maximum" = ~ max(.data$WKHP),
            "Median" = ~ median(.data$WKHP),
            "Mean" = ~ mean(.data$WKHP))            
            # "Mean" = ~ mean(.data$WAGP),
            # "Mean (sd)" = ~ qwraps2::mean_sd(.data$WKHP))
       )
### Overall
whole <- summary_table(ss16ppr, summary3)
kable(whole) %>%
  kable_styling(bootstrap_options = "striped")
```


<!-- We were asked to perform an exploratory data analysis (EDA) based on American Community Survey, and our research questions are: 1. In Puerto Rico, how does earnings vary by educational attainment? 2. How does earnigs vary by how well English is spoken?  -->

<!-- Based on these two questions, we needed several attributes from the survey, which are ss16ppr$RT, ss16ppr$RAC1P, ss16ppr$RAC2P, ss16ppr$RAC3P, ss16ppr$ENG, ss16ppr$DREM, ss16ppr$SCHL. In detail, we looked into race diversity, marital status, ability to speak English, cognitive difficulty, educational attainment, and gender in Puerto Rico. -->


### Ethicity: Hispanic

```{r}
kable(ss16ppr %>% group_by(RACHISP) %>% summarise(Count = n())) %>%
  kable_styling(bootstrap_options = "striped")
```

### Race: White

In Puerto Rico, the majority of people identify themselves as white.  Minority races including American Indian, Alaska Native, Asian, Native Hawaiian and Other Pacific Islander can be eliminated.

```{r}
kable(ss16ppr %>% group_by(RACWHT) %>% summarise(Count = n())) %>%
  kable_styling(bootstrap_options = "striped")
```

### Race: Black

```{r}
kable(ss16ppr %>% group_by(RACBLK) %>% summarise(Count = n())) %>%
  kable_styling(bootstrap_options = "striped")
```

### Marital status

```{r}
kable(ss16ppr %>%  group_by(MAR) %>% summarise(Count = n())) %>%
  kable_styling(bootstrap_options = "striped") 

kable(ss16ppr %>% group_by(MAR1) %>% summarise(Count = n())) %>%
  kable_styling(bootstrap_options = "striped")
```

## Ability to speak English

```{r}
kable(ss16ppr %>%  group_by(ENG) %>% summarise(n())) %>%
  kable_styling(bootstrap_options = "striped")
```

# Educational attainment

+ We will categorize people into seven education levels
    + HS diploma or equivalent
    + Some college
    + Associate's degree
    + Bachelor's degree
    + Professional degree
    + Doctorate degree
    
```{r}
kable(ss16ppr %>%  group_by(SCHL) %>% summarise(Count = n())) %>%
  kable_styling(bootstrap_options = "striped")
```

## Sex 

```{r}
#code to decode SEX
ss16ppr <- mutate(ss16ppr, SEX = derivedFactor(
  "Male" = SEX == 1,
  "Female" = SEX == 2,
  method ="first",
  .default = "NA"
))
kable(ss16ppr %>%  group_by(SEX) %>% summarise(Count = n())) %>%
  kable_styling(bootstrap_options = "striped")
```

## Task 3:

Generate and examine histograms for quantitative variables of interest

### Wage

```{r}
ggplot(ss16ppr,  aes(WAGP)) +
  geom_histogram(bins = 15)  +
  theme_minimal() + 
  theme(axis.text.x = element_text(vjust = 1, 
                                   size = 12, hjust = 1))
```

### Age
 
```{r}
ggplot(ss16ppr,  aes(AGEP)) +
  geom_histogram(stat="count")  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1))
```

### Hours worked

```{r}
ggplot(ss16ppr,  aes(WKHP)) +
  geom_histogram(stat="count")  +
  theme_minimal() 
```

## Task 4:

Generate and examine bar charts/graphs for qualitative variables of interest

### Race

```{r}
ggplot(ss16ppr,  aes(RAC1P)) +
  geom_bar()  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))
```

### White

```{r}
ggplot(ss16ppr,  aes(RACWHT)) +
  geom_bar()  +
  theme_minimal()
```

### Black

```{r}
ggplot(ss16ppr,  aes(RACBLK)) +
  geom_bar()  +
  theme_minimal()
```

### Hispanic

```{r}
ggplot(ss16ppr,  aes(RACHISP)) +
  geom_bar()  +
  theme_minimal()
```

### Marital Status

```{r}
ggplot(ss16ppr,  aes(MAR)) +
  geom_bar()  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))
```

```{r}
ggplot(ss16ppr,  aes(MAR1)) +
  geom_bar()  +
  theme_minimal()  
```

## Ability to speak English

About half of population does not speak English well

```{r}
ggplot(ss16ppr,  aes(ENG)) +
  geom_bar()  +
  theme_minimal()  
```

## Educational attainment

```{r}
ggplot(ss16ppr,  aes(SCHL)) +
  geom_bar()  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))
```

## Sex 

Gender is nearly equalized in Puerto Rico

```{r}
ggplot(ss16ppr,  aes(SEX)) +
  geom_bar()  +
  theme_minimal()
```

### Work week

```{r}
ggplot(ss16ppr,  aes(WKW)) +
  geom_histogram(stat="count")  +
  theme_minimal()
```

Task 5:

Generate and examine cross tabulations, scatterplots, and/or correlation coefficients of interest

```{r}
ss16ppr <- filter(ss16ppr, WAGP != max(ss16ppr$WAGP))
ggplot(ss16ppr,  aes(AGEP, WAGP)) +
  geom_point() 
```

```{r}
cor(ss16ppr$WAGP, ss16ppr$AGEP)
```

The correlation of .23 for age and earnings indicates a very weak relationship. Age is neither a primary reason for differences in earnings, nor a clear proxy for some other variable.

```{r}
ggplot(ss16ppr,  aes(WKHP, WAGP)) +
  geom_point()  
```

```{r}
cor(ss16ppr$WAGP, ss16ppr$WKHP)
```

The correlation of .18 for earnings and work hours is also very weak. No doubt it would be strong if the data were not filtered to those working more than 35 hours per week. Interestingly, earnings appear to drop for those working more than 60 hours per week.

```{r}
### By RACHISP
by_RACHISP <- summary_table(dplyr::group_by(ss16ppr, RACHISP), summary1)
kable(by_RACHISP) %>%
  kable_styling(bootstrap_options = "striped")
```

```{r}
### By RACWHT
by_RACWHT <- summary_table(dplyr::group_by(ss16ppr, RACWHT), summary1)
kable(by_RACWHT) %>%
  kable_styling(bootstrap_options = "striped")
```

```{r}
### By RACBLK
by_RACBLK <- summary_table(dplyr::group_by(ss16ppr, RACBLK), summary1)
kable(by_RACBLK) %>%
  kable_styling(bootstrap_options = "striped")
```
