---
title: "Labor Market Analysis"
author: "Marjorie Blanco, Joe Thomson, Haodi Tu"
subtitle: Exploratory Data Analysis
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r, echo=FALSE}
# Course: ECON 5330
# Title: EDA Assignment
# Date: Jan 22, 2019
# Author: Haodi Tu, Majorie Blanco, Joseph Thomson
```

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(readr)
library(ggplot2)
library(mosaic)
library(dplyr)
library(kableExtra)
library(qwraps2)
```


```{r}
fun_mean <- function(x){
  return(data.frame(y=mean(x),label=round(mean(x,na.rm=T), 2)))}
```

## Research Questions and Geographic area 

## Data

<!-- https://cran.r-project.org/web/packages/qwraps2/vignettes/summary-statistics.html -->

```{r}
ss16ppr <- read_csv("csv_ppr/ss16ppr.csv")
#kable(ss16ppr)
```

We used data from the 2016 ACS for Puerto Rico to examine earnings gaps between individuals with different education levels. Our research questions are:
1) How do earnings vary by education level?
2) How does the premium for education vary by gender?

The 2016 ACS is a nationally representative sample of `r nrow(ss16ppr)`  households with a response rate of 97.6% (TBD). The household survey includes questions pertaining to each household member’s demographic characteristics and labor market activity. Our analysis relies on data collected on individuals living in households but not on those living in group quarters.

We restrict our sample to these three racial groups: White, Black and Other. In addition, given our goal of examining earnings differences by ability to speak English and the reporting of earnings in the ACS on an annual basis (wages, salary, commissions, bonuses, tips, and self-employment income during the past 12 months), we restrict our sample to full-time year-round (FTYR) workers. We define FTYR workers as individuals who report positive earnings over the past year, who worked at least 40 of the past 52 weeks, and who worked at least 35 hours per week in a usual work week over this period. 

```{r}
#select columns of interest
ss16ppr <- ss16ppr %>% select(PERNP, SEX, AGEP, RAC1P, RACWHT, RACBLK, HISP, MAR, SCHL, ENG, COW, WKW, WKHP)
```

## Sample selection criteria

+ Restricted age between 18 to 64 years old. The reason is that we only need to investage people at working age.
+ Categorized people in seven education levels
    + Eliminated people who do not have education (before high school)
    + HS diploma or equivalent
    + Some college
    + Associate's degree
    + Bachelor's degree
    + Professional degree
    + Doctorate degree
+ Removed NAs and zero in wage segment
    + Removed outliers
+ Restricted class of worker to exclude those working without pay and unemployed
+ Selected peoplewho worked at least 35 hours/week


```{r}
ss16ppr$WKHP <- as.numeric(ss16ppr$WKHP)
ss16ppr$PERNP <- as.numeric(ss16ppr$PERNP)
ss16ppr$AGEP <- as.numeric(ss16ppr$AGEP)


#Filters
# AGE: Only between 18 and 64 (inclusive) years of age
ss16ppr <- ss16ppr %>% filter(AGEP >= 18, AGEP <= 64)

# SCHL: High school and above
#       Combine: 16/17 for High school and 18/19 for Some college
ss16ppr <- ss16ppr %>% filter(SCHL >= 16) %>% mutate(SCHL = ifelse(SCHL == 17, 16, SCHL)) %>% mutate(SCHL = ifelse(SCHL == 19, 18, SCHL))

# WKW: Only between 50 and 40 hours
# 1 .50 to 52 weeks worked during past 12 months
# 2 .48 to 49 weeks worked during past 12 months
# 3 .40 to 47 weeks worked during past 12 months
ss16ppr <- ss16ppr %>% filter(between(WKW, 1,3))

# Remove NAs and zero in PERNP
ss16ppr <- ss16ppr %>% filter(!is.na(PERNP))
ss16ppr <- ss16ppr %>% filter(PERNP != 0) 
# Remove outlier
#ss16ppr <-ss16ppr %>% filter(PERNP != max(ss16ppr$PERNP))

# Restirct class of worker to exclude those working without pay and unemployed
# Class of worker
# b N/A (less than 16 years old/NILF who last worked more than 5 years ago or never worked)
# 1 Employee of a private for-profit company or business, or of an individual, for wages,
#    salary, or commissions
# 2 Employee of a private not-for-profit, tax-exempt, or charitable organization
# 3 Local government employee (city, county, etc.)
# 4 State government employee
# 5 Federal government employee
# 6 Self-employed in own not incorporated business, professional practice, or farm
# 7 Self-employed in own incorporated business, professional practice or farm
# 8 Working without pay in family business or farm
# 9 Unemployed and last worked 5 years ago or earlier or never worked
ss16ppr <- ss16ppr %>% filter(between(COW, 1, 7))


# WKHP: worked at least 35 hr per week in a usual work week over this period. 
ss16ppr <- ss16ppr %>% filter(WKHP >= 35)

ss16ppr <- ss16ppr %>% filter(PERNP < 159000) %>% filter(PERNP >= 10000)
```


```{r}
# Decode variables of interest

ss16ppr  <- ss16ppr %>% mutate(Female = ifelse(SEX == 2, 1, 0))


ss16ppr  <- ss16ppr %>% mutate(RACHISP = ifelse(HISP == "01", "Yes", "No"))

ss16ppr  <- ss16ppr %>% mutate(RaceBlack = RACBLK)
ss16ppr  <- ss16ppr %>% mutate(RaceOther = ifelse(RACBLK == 1 | RACWHT == 1, 0, 1))


ss16ppr  <- ss16ppr %>% mutate(RACOTHER = ifelse(RACBLK == 1 | RACWHT == 1, "No", "Yes"))
ss16ppr  <- ss16ppr %>% mutate(RACWHT = ifelse(RACWHT == 1, "Yes", "No"))
ss16ppr  <- ss16ppr %>% mutate(RACBLK = ifelse(RACBLK == 1, "Yes", "No"))

ss16ppr <- ss16ppr %>% mutate(MAR1 = ifelse(MAR == 1, "Yes", "No"))

#code to decode RAC1P
ss16ppr <- mutate(ss16ppr, RAC1P = derivedFactor(
  "White" = RAC1P == 1,
  "Black or African American" = RAC1P == 2,
  "American Indian" = RAC1P == 3,
  "Alaska Native" = RAC1P == 4,
  "American Indian and Alaska Native" = RAC1P == 5,
  "Asian" = RAC1P == 6,
  "Native Hawaiian and Other Pacific Islander" = RAC1P == 7,
  "Some Other " = RAC1P == 8,
  "Two or More Races" = RAC1P == 9,
  method ="first",
  .default = "NA"
))


ss16ppr  <- ss16ppr %>% mutate(Divorced = ifelse(MAR == 3 | MAR == 4, 1, 0))
ss16ppr  <- ss16ppr %>% mutate(NeverMarried = ifelse(MAR == 5, 1, 0))


#code to decode MAR
ss16ppr <- mutate(ss16ppr, MAR = derivedFactor(
  "Married" = MAR == 1,
  "Widowed" = MAR == 2,
  "Divorced" = MAR == 3,
  "Separated" = MAR == 4,
  "Never married" = MAR == 5,
  method ="first",
  .default = "NA"
))

#Note: 16 .Regular high school diploma

ss16ppr  <- ss16ppr %>% mutate(SomeCollege = ifelse(SCHL == 18, 1, 0))
ss16ppr  <- ss16ppr %>% mutate(Associate = ifelse(SCHL == 20, 1, 0))
ss16ppr  <- ss16ppr %>% mutate(Bachelor = ifelse(SCHL == 21, 1, 0))
ss16ppr  <- ss16ppr %>% mutate(Master = ifelse(SCHL == 22, 1, 0))
ss16ppr  <- ss16ppr %>% mutate(Professional = ifelse(SCHL == 23, 1, 0))
ss16ppr  <- ss16ppr %>% mutate(Doctoral = ifelse(SCHL == 24, 1, 0))


#code to decode SCHL
ss16ppr <- mutate(ss16ppr, SCHL = derivedFactor(
  "High school diploma" = SCHL == "16",
  "Some college" = SCHL == "18",
  "Associate's degree" = SCHL == "20",
  "Bachelor's degree" = SCHL == "21",
  "Master's degree" = SCHL == "22",
  "Professional degree" = SCHL == "23",
  "Doctorate degree" = SCHL == "24",
  method ="first",
  .default = "NA"
))

# 1 .50 to 52 weeks worked during past 12 months
# 2 .48 to 49 weeks worked during past 12 months
# 3 .40 to 47 weeks worked during past 12 months

#code to decode WKW
ss16ppr <- mutate(ss16ppr, WKW = derivedFactor(
  "50 to 52" = WKW == 1,
  "48 to 49" = WKW == 2,
  "40 to 47" = WKW == 3,
  method ="first",
  .default = "NA"
))
```

```{r}
#head(ss16ppr %>% arrange(PERNP), 90)

#head(ss16ppr %>% arrange(desc(WKHP)), 90)


ggplot(ss16ppr %>% filter(COW == 6),  aes(PERNP)) +
  geom_histogram(bins = 15)  +
  theme_minimal() + 
  theme(axis.text.x = element_text(vjust = 1, 
                                   size = 12, hjust = 1)) +
  xlab("")
```


We used data from the 2016 ACS for Puerto Rico to examine wage gaps between individuals with different education levels. Our research questions are:
1) How do earnings vary by education level?
2) How does the premium for education vary by gender?
The 2016 ACS is a nationally representative sample of `r nrow(ss16ppr)`. The household survey includes questions pertaining to each household member’s demographic characteristics and labor market activity. 

We restrict our sample to these three racial groups: White, Black and Other. In addition, given our goal of examining earning differences by gender and marital status and the reporting of earnings in the ACS on an annual basis (wages, salary, commissions, bonuses, tips, and self-employment income during the past 12 months), we restrict our sample to full-time year-round (FTYR) workers. We define FTYR workers as individuals who report positive earnings over the past year, who worked at least 40 of the past 52 weeks, and who worked at least 35 hours per week in a usual work week over this period. 

## EDA Insights: 

For our exploratory analysis we looked at population breakdowns by education, age,  marital status, gender, race, earnings, and work hours. We applied filters on education (HS diploma or above), age (18-64), and work hours (>35/week). 

An earnings histogram identified a default maximum amount of earnings (189k) which we also filtered out of the data. The earning distribution is progressive above the median, but drops off sharply below the median, likely indicating the presence of a minimum wage. The correlation between age and earnings is very weak (.23). Likewise, earnings is very weakly correlated with hours worked among those who work more than 35 hours per week. However, white individuals appear to have an earnings premium over other races, and both married and divorced individuals appear to have an earnings premium over those who have never been married. Given that the correlation between age and earnings was weak, this may be due to other qualitative factors possessed by those who get married.  Married was recategoried to married and not married.

Earnings appear positively correlated with how well people speak English, as well as with higher levels of education. It should be noted that, on the island, fluency in the English language is neither a requirement, nor really needed. As a Latin American destination, the predominant language spoken and used (e.g. street signs, day-to-day communications) is Spanish, with English only coming into use in the tourism industry, or in those industies or companies imported from the mainland. This may help to explain the earning difference, as those are likely to pay more than local businesses (note that tourism is the largest industry on the island and the source of most of the island's GDP). Men also appear to earn a small premium over women.


The age distribution of full time workers is skewed towards older adults, possibly indicating that younger workers have trouble finding full-time work, wait to enter the workforce, or are leaving the territory.

## Task 1: 

Examine the first 10 or 20 observations (rows of data) corresponding to variables of interest (columns) and compare the observed values to the data dictionary for person records.

```{r}
kable(head(ss16ppr, 20)) %>%
  kable_styling(bootstrap_options = "striped")
```

## Task 2:

Compute and examine descriptive statistics including the minimum, maximum, mean, and median for quantitative variables of interest

### Total person's earnings

```{r}
ss16ppr$PERNP <- as.numeric(ss16ppr$PERNP)
summary1 <-
  list("Wages or salary income past 12 months" =
         list("Minimum" = ~ min(.data$WAGP),
              "Maximum" = ~ max(.data$WAGP),
              "Median" = ~ median(.data$WAGP),
              "Mean" = ~ mean(.data$WAGP))
  list("Total person's earnings" =
         list("Minimum" = ~ min(.data$PERNP),
              "Maximum" = ~ max(.data$PERNP),
              "Median" = ~ median(.data$PERNP),
              "Mean" = ~ mean(.data$PERNP))
  )

### Overall
whole <- summary_table(ss16ppr, summary1)
kable(whole) %>%
  kable_styling(bootstrap_options = "striped")
```

### Age

```{r}
summary2 <-
  list("Age" =
         list("Minimum" = ~ min(.data$AGEP),
              "Maximum" = ~ max(.data$AGEP),
              "Median" = ~ median(.data$AGEP),
              "Mean" = ~ mean(.data$WAGP))
)

              "Mean" = ~ mean(.data$PERNP))
)

### Overall
whole <- summary_table(ss16ppr, summary2)
kable(whole) %>%
  kable_styling(bootstrap_options = "striped")
```

### Hours worked 

```{r}
summary3 <-
  list("Usual hours worked per week past 12 months" =
         list("Minimum" = ~ min(.data$WKHP),
              "Maximum" = ~ max(.data$WKHP),
              "Median" = ~ median(.data$WKHP),
              "Mean" = ~ mean(.data$WKHP))
  )
### Overall
whole <- summary_table(ss16ppr, summary3)
kable(whole) %>%
  kable_styling(bootstrap_options = "striped")
```



### Ethicity: Hispanic

```{r}
kable(ss16ppr %>% group_by(RACHISP) %>% summarise(Count = n())) %>%
  kable_styling(bootstrap_options = "striped")
```

### Race: White

In Puerto Rico, the majority of people identify themselves as white.  Minority races including American Indian, Alaska Native, Asian, Native Hawaiian and Other Pacific Islander can be eliminated.

```{r}
kable(ss16ppr %>% group_by(RACWHT) %>% summarise(Count = n())) %>%
  kable_styling(bootstrap_options = "striped")
```

### Race: Black

```{r}
kable(ss16ppr %>% group_by(RACBLK) %>% summarise(Count = n())) %>%
  kable_styling(bootstrap_options = "striped")
```

### Race: Other

```{r}
kable(ss16ppr %>% group_by(RACOTHER) %>% summarise(Count = n())) %>%
  kable_styling(bootstrap_options = "striped")
```

### Marital status

```{r}
kable(ss16ppr %>%  group_by(MAR) %>% summarise(Count = n())) %>%
  kable_styling(bootstrap_options = "striped") 

kable(ss16ppr %>% group_by(MAR1) %>% summarise(Count = n())) %>%
  kable_styling(bootstrap_options = "striped")
```

# Educational attainment

+ We will categorize people into seven education levels
+ HS diploma or equivalent
+ Some college
+ Associate's degree
+ Bachelor's degree
+ Professional degree
+ Doctorate degree

```{r}
kable(ss16ppr %>%  group_by(SCHL) %>% summarise(Count = n())) %>%
  kable_styling(bootstrap_options = "striped")
```

## Sex 

```{r}
#code to decode SEX
ss16ppr <- mutate(ss16ppr, SEX = derivedFactor(
  "Male" = SEX == 1,
  "Female" = SEX == 2,
  method ="first",
  .default = "NA"
))
kable(ss16ppr %>%  group_by(SEX) %>% summarise(Count = n())) %>%
  kable_styling(bootstrap_options = "striped")
```

## Task 3:

Generate and examine histograms for quantitative variables of interest

### Total person's earnings

+ Right skewed wage histogram
    + Only a few have high wage on right end;
    + Median wage is around $24,000;
    + Average is around 28,000, a little bit higher than median.
    

```{r}
ggplot(ss16ppr,  aes(PERNP)) +
  geom_histogram(bins = 15)  +
  theme_minimal() + 
  theme(axis.text.x = element_text(vjust = 1, 
                                   size = 12, hjust = 1))
```

### Age

+ Bell shape alike for age histogram 
  + median and mean are close (around 43 years old)


```{r}
ggplot(ss16ppr,  aes(AGEP)) +
  geom_histogram(stat="count")  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1))
```

### Hours worked

+ 40 hr per week is the standard weekly working hour in Puerto Rico

```{r}
ggplot(ss16ppr,  aes(WKHP)) +
  geom_histogram(stat="count")  +
  theme_minimal() 
```

## Task 4:

Generate and examine bar charts/graphs for qualitative variables of interest

### Race

```{r}
ggplot(ss16ppr,  aes(RAC1P)) +
  geom_bar()  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))

ggplot(ss16ppr,  aes(RAC1P, PERNP)) +
  geom_boxplot() +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  stat_summary(fun.data = fun_mean, geom="text", vjust=-0.7, size=3)
```

### White

```{r}
ggplot(ss16ppr,  aes(RACWHT)) +
  geom_bar()  +
  theme_minimal()

ggplot(ss16ppr,  aes(RACWHT, PERNP)) +
  geom_boxplot() +
  theme_minimal() +
  stat_summary(fun.data = fun_mean, geom="text", vjust=-0.7, size=3)
```

### Black

```{r}
ggplot(ss16ppr,  aes(RACBLK)) +
  geom_bar()  +
  theme_minimal() +
  xlab("")

ggplot(ss16ppr,  aes(RACBLK, PERNP)) +
  geom_boxplot() +
  theme_minimal() +
  stat_summary(fun.data = fun_mean, geom="text", vjust=-0.7, size=3) +
  xlab("")
```

### Race: Other

```{r}
ggplot(ss16ppr,  aes(RACOTHER)) +
  geom_bar()  +
  theme_minimal()

ggplot(ss16ppr,  aes(RACOTHER, PERNP)) +
  geom_boxplot() +
  theme_minimal() +
  stat_summary(fun.data = fun_mean, geom="text", vjust=-0.7, size=3)
```

### Marital Status

```{r}
ggplot(ss16ppr,  aes(MAR)) +
  geom_bar()  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))

ggplot(ss16ppr,  aes(MAR, PERNP)) +
  geom_boxplot() +
  theme_minimal() +
  stat_summary(fun.data = fun_mean, geom="text", vjust=-0.7, size=3)
```

```{r}
ggplot(ss16ppr,  aes(MAR1)) +
  geom_bar()  +
  theme_minimal() +
  stat_summary(fun.data = fun_mean, geom="text", vjust=-0.7, size=3)
```

### Ability to speak English


```{r}
ggplot(ss16ppr,  aes(ENG)) +
  geom_bar()  +
  theme_minimal()  
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1)) 


ggplot(ss16ppr,  aes(MAR1, PERNP)) +
  geom_boxplot() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1)) +
  stat_summary(fun.data = fun_mean, geom="text", vjust=-0.7, size=3)
```

### Educational attainment

```{r}
ggplot(ss16ppr,  aes(SCHL)) +
  geom_bar()  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))

ggplot(ss16ppr,  aes(SCHL, PERNP)) +
  geom_boxplot() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1)) +
  stat_summary(fun.data = fun_mean, geom="text", vjust=-0.7, size=3)
```

### Sex 


```{r}
ggplot(ss16ppr,  aes(SEX)) +
  geom_bar()  +
  theme_minimal()

ggplot(ss16ppr,  aes(SEX, PERNP)) +
  geom_boxplot() +
  theme_minimal() +
  stat_summary(fun.data = fun_mean, geom="text", vjust=-0.7, size=3)
```

### Work week

```{r}
ggplot(ss16ppr,  aes(WKW)) +
  geom_bar()  +
  theme_minimal()

ggplot(ss16ppr,  aes(WKW, PERNP)) +
  geom_boxplot() +
  theme_minimal() +
  stat_summary(fun.data = fun_mean, geom="text", vjust=-0.7, size=3)
```

##Task 5:

Generate and examine cross tabulations, scatterplots, and/or correlation coefficients of interest

<<<<<<< HEAD
### Correlation between age and wage

The correlation of `r round(cor(ss16ppr$WAGP, ss16ppr$AGEP),2)` for age and earnings indicates a very weak relationship. Age is neither a primary reason for differences in earnings, nor a clear proxy for some other variable.

+ Correlation is 0.2348 
    + Indicates there is a positive relaitonship between age and wage
    
```{r}
ss16ppr <- filter(ss16ppr, WAGP != max(ss16ppr$WAGP))
ggplot(ss16ppr,  aes(AGEP, WAGP)) +
  geom_point() 
```
=======
### Age vs. Total person's earnings 
>>>>>>> 4986b7974727e968702a7e6d14d89e59677ceafa

```{r}
ss16ppr <- filter(ss16ppr, PERNP != max(ss16ppr$PERNP))
ggplot(ss16ppr,  aes(AGEP, PERNP)) +
  geom_point() +
  xlab("")
```

<<<<<<< HEAD


### Correlation between work hour and wage

The correlation of `r round(cor(ss16ppr$WAGP, ss16ppr$WKHP),2)` for earnings and work hours is also very weak. No doubt it would be strong if the data were not filtered to those working more than 35 hours per week. Interestingly, earnings appear to drop for those working more than 60 hours per week.

+ Correlation is 0.1829 
    + Indicates there is a positive relaitonship between work hour and wage
    + The relation is weaker than correlation between age and wage

=======
The correlation of `r round(cor(ss16ppr$PERNP, ss16ppr$AGEP),2)` for age and earnings indicates a very weak relationship. Age is neither a primary reason for differences in earnings, nor a clear proxy for some other variable.

### Work hours vs. Total person's earnings 
>>>>>>> 4986b7974727e968702a7e6d14d89e59677ceafa

```{r}
ggplot(ss16ppr,  aes(WKHP, PERNP)) +
  geom_point()  +
  xlab("")
```

<<<<<<< HEAD


### Wage comparison by race

+ Hispanic people in Puerto Rico make less money then other races on average
    + $6000 less by median
    + approximately $1500 less on average
=======
The correlation of `r round(cor(ss16ppr$PERNP, ss16ppr$WKHP),2)` for earnings and work hours is also very weak. No doubt it would be strong if the data were not filtered to those working more than 35 hours per week. Interestingly, earnings appear to drop for those working more than 60 hours per week.
>>>>>>> 4986b7974727e968702a7e6d14d89e59677ceafa

```{r}
### By RACHISP
by_RACHISP <- summary_table(dplyr::group_by(ss16ppr, RACHISP), summary1)
kable(by_RACHISP) %>%
  kable_styling(bootstrap_options = "striped")
```


+ White people in Puerto Rico make more money than other races on average 
    + $1250 more by median
    + approximately $3000 more on average
    
```{r}
### By RACWHT
by_RACWHT <- summary_table(dplyr::group_by(ss16ppr, RACWHT), summary1)
kable(by_RACWHT) %>%
  kable_styling(bootstrap_options = "striped")
```

+ Black people in Puerto Rico make less money than other races on average 
    + $1150 less by median
    + approximately $2000 less on average

```{r}
### By RACBLK
by_RACBLK <- summary_table(dplyr::group_by(ss16ppr, RACBLK), summary1)
kable(by_RACBLK) %>%
  kable_styling(bootstrap_options = "striped")
```

<<<<<<< HEAD
+ Other race people in Puerto Rico make less money than other races on average 
    + $1000 less by median
    + approximately $2000 less on average



```{r}
### By RACOTHER

by_RACOTHER <- summary_table(dplyr::group_by(ss16ppr, RACOTHER), summary1)
kable(by_RACOTHER) %>%
  kable_styling(bootstrap_options = "striped")
```
<<<<<<< HEAD
=======

### Marital Status

```{r}
### By MAR1
by_MAR <- summary_table(dplyr::group_by(ss16ppr, MAR1), summary1)
kable(by_MAR) %>%
  kable_styling(bootstrap_options = "striped")
```

### Educational attainment

```{r}
### By SCHL
by_SCHL <- summary_table(dplyr::group_by(ss16ppr, SCHL), summary1)
kable(by_SCHL) %>%
  kable_styling(bootstrap_options = "striped")
```


## First Model: 
$$Earning = \beta_0 + Divorced * \beta_1 + NeverMarried * \beta_2 + Female * \beta_3 + RaceBlack * \beta_4 + RaceOther * \beta_5 + SomeCollege * \beta_6 + Associate * \beta_7 + Bachelor * \beta_8 + Master * \beta_9 + Professional * \beta_10 + Doctoral * \beta_11 + Age * \beta_12$$

```{r}
# REFRENCE: HIGH SCHOOL
model_1 <- lm(PERNP ~ Divorced + NeverMarried + Female + RaceBlack + RaceOther + SomeCollege + Associate + Bachelor + Master + Professional + Doctoral + AGEP, data = ss16ppr)

summary(model_1)
```

+ Coefficients Explaination 
    + Holding gender, race, education and age constant, married or widowed people makes $1142 more than people who divorced or separated on average.
    + Holding gender, race, education and age constant, married or widowed people makes $2948 more than people who never married on average.
    + Holding marriage, race, education and age constant, male makes $4821 more than female on average.
    + Holding marriage, gender, education and age constant, White makes $2195 less than Black on average.
    + Holding marriage, gender, education and age constant, White makes $1 more than Other race on average.
    + Holding marriage, gender, race and age constant, people have high school education makes $4327 less than people have some college education on average.
    + Holding marriage, gender, race and age constant, people have high school education makes $4222 less than people have associate education on average.
    + Holding marriage, gender, race and age constant, people have high school education makes $12374 less than people have Bachelor degree on average.
    + Holding marriage, gender, race and age constant, people have high school education makes $17826 less than people have Master degree on average.
    + Holding marriage, gender, race and age constant, people have high school education makes $28089 less than people have Professional education on average.
    + Holding marriage, gender, race and age constant, people have high school education makes $35671 less than people have Doctoral degree on average.
    + Holding marriage, gender, race and education constant, people make $287 more as age increases on average between the age of 18 to 64.
    
    


## Second Model:
$$Earning = \beta_0 + Female * \beta_1 + SomeCollege * \beta_2 + Associate * \beta_3 + Bachelor * \beta_4 + Master * \beta_5 + Professional * \beta_6 + Doctoral * \beta_7$$


```{r}
model_2 <-lm(PERNP ~ Female + SomeCollege + Associate + Bachelor + Master + Professional + Doctoral, data = ss16ppr)
summary(model_2)
```

+ Coefficients Explaination
    + Holding education constant, male makes $4656 more than female on average.
    + Holding gender constant, people have high school education makes $3340 less than people have some college education on average.
    + Holding gender constant, people have high school education makes $4001 less than people have associate education on average.
    + Holding gender constant, people have high school education makes $12229 less than people have Bachelor degree on average.
    + Holding gender constant, people have high school education makes $17934 less than people have Master degree on average.
    + Holding gender constant, people have high school education makes $28336 less than people have Professional education on average.
    + Holding gender constant, people have high school education makes $37602 less than people have Doctoral degree on average.
    
    
    
    
    
    
    
    
    
    
