---
title: "Lending Mortage Analysis"
author: "Marjorie Blanco, Joe Thompson, Haodi Tu"
output:
  pdf_document: default
  html_document: default
subtitle: Exploratory Data Analysis
geometry: left=2cm,right=3cm,top=2cm,bottom=2cm
---


```{r warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(scipen=999)
library(readr)
library(ggplot2)
library(readxl)
library(kableExtra)
library(dplyr)
library(tidyverse)
library(purrr)
library(ggthemes)
library(aod)
library(lmtest)
library(survey)
library(caret)
library(pscl)
library(survey)

```

```{r}
data <- read_excel("MLD/MLD Data File.xls")

#data <- read_excel("MLD Data File.xls")

data <- tibble::rowid_to_column(data, "ID")
data$RACE <- ifelse(data$BLACK == 1, "non-Hispanic Black", ifelse(data$HISPAN == 1, "Hispanic", "non-Hispanic White"))

data  <- data %>% mutate(MARRIED = ifelse(MARRIED == 1, "Yes", ifelse(MARRIED == 0, "No", "Unknown")))
data  <- data %>% mutate(BLACK = ifelse(BLACK == 1, "Yes", ifelse(BLACK == 0, "No", "Unknown")))
data  <- data %>% mutate(HISPAN = ifelse(HISPAN == 1, "Yes", ifelse(HISPAN == 0, "No", "Unknown")))
data  <- data %>% mutate(APPROVE = ifelse(APPROVE == 1, "Yes", ifelse(APPROVE == 0, "No", "Unknown")))
data  <- data %>% mutate(MALE = ifelse(MALE == 1, "Yes", ifelse(MALE == 0, "No", "Unknown")))
data$LOANPRC <- data$LOANPRC * 100

data$MARRIED <- as.factor(data$MARRIED)
data$MALE <- as.factor(data$MALE)
data$APPROVE <- as.factor(data$APPROVE)
data$GDLIN <- as.factor(data$GDLIN)
data$BLACK <- as.factor(data$BLACK)
data$HISPAN <- as.factor(data$HISPAN)

data$RACE <- as.factor(data$RACE)
data1 <- data

colnames(data1) <- c("ID", "Married", "Meet credit history guidelines",  "Other obligations as a percent of total income", "non-Hispanic Black",  "Hispanic", "Male", "Mortgage loan approved", "Loan amount/purchase price", "Race")
```

The data set contains `r nrow(data)` records.  The overall descriptive statistics:

```{r}
summary(data1[,2:10])
```

Descriptive statistics by Race:

```{r}
data %>% select(-c(ID, BLACK, HISPAN)) %>% split(.$RACE) %>% map(summary) 
```

```{r}
ggplot(data,  aes(OBRAT)) +
  geom_histogram(binwidth = 1)  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 0, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Other obligations as a percent of total income") +
  facet_grid(RACE ~ ., scales="free")



data1 <- data

data1$OBRAT<-cut(data1$OBRAT, seq(0,95,1))
data1$LOANPRC<-cut(data1$LOANPRC, seq(0,260,5))


ggplot() +
  geom_bar(data = data1 %>% filter(RACE == "Hispanic"),  aes(x = OBRAT, fill = APPROVE), position = "fill")  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Other obligations as a percent of total income") +
  scale_y_continuous("Percent") +
  scale_fill_manual(values = c("red", "green")) +
  facet_grid(RACE ~ ., scales="free") +
  theme(legend.position="bottom")

ggplot() +
  geom_bar(data = data1 %>% filter(RACE == "non-Hispanic Black"),  aes(x = OBRAT, fill = APPROVE), position = "fill")  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Other obligations as a percent of total income") +
  scale_y_continuous("Percent") +
  scale_fill_manual(values = c("red", "green")) +
  facet_grid(RACE ~ ., scales="free") +
  theme(legend.position="bottom")

ggplot() +
  geom_bar(data = data1 %>% filter(RACE == "non-Hispanic White"),  aes(x = OBRAT, fill = APPROVE), position = "fill")  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Other obligations as a percent of total income") +
  scale_y_continuous("Percent") +
  scale_fill_manual(values = c("red", "green")) +
  facet_grid(RACE ~ ., scales="free") +
  theme(legend.position="bottom")

ggplot() +
  geom_bar(data = data1 %>% filter(MALE == "Yes"),  aes(x = OBRAT, fill = APPROVE), position = "fill")  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Other obligations as a percent of total income") +
  scale_y_continuous("Percent") +
  scale_fill_manual(values = c("red", "green")) +
  facet_grid(MALE ~ ., scales="free", labeller = as_labeller(c(Yes = "Male"))) +
  theme(legend.position="bottom")

ggplot() +
  geom_bar(data = data1 %>% filter(MALE == "No"),  aes(x = OBRAT, fill = APPROVE), position = "fill")  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Other obligations as a percent of total income") +
  scale_y_continuous("Percent") +
  scale_fill_manual(values = c("red", "green")) +
  facet_grid(MALE ~ ., scales="free", labeller = as_labeller(c(No = "Female"))) +
  theme(legend.position="bottom") 


ggplot() +
  geom_bar(data = data1 %>% filter(MARRIED == "Yes"),  aes(x = OBRAT, fill = APPROVE), position = "fill")  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Other obligations as a percent of total income") +
  scale_y_continuous("Percent") +
  scale_fill_manual(values = c("red", "green")) +
  facet_grid(MARRIED ~ ., scales="free", labeller = as_labeller(c(Yes = "Married"))) +
  theme(legend.position="bottom")

ggplot() +
  geom_bar(data = data1 %>% filter(MARRIED == "No"),  aes(x = OBRAT, fill = APPROVE), position = "fill")  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Other obligations as a percent of total income") +
  scale_y_continuous("Percent") +
  scale_fill_manual(values = c("red", "green")) +
  facet_grid(MARRIED ~ ., scales="free", labeller = as_labeller(c(No = "Not Married"))) +
  theme(legend.position="bottom")
```

```{r}
ggplot() +
  geom_bar(data = data1 %>% filter(RACE == "Hispanic"),  aes(x = LOANPRC, fill = APPROVE), position = "fill")  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Loan amount/purchase price %") +
  scale_y_continuous("Percent") +
  scale_fill_manual(values = c("red", "green")) +
  facet_grid(RACE ~ ., scales="free") +
  theme(legend.position="bottom") 

ggplot() +
  geom_bar(data = data1 %>% filter(RACE == "non-Hispanic Black"),  aes(x = LOANPRC, fill = APPROVE), position = "fill")  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Loan amount/purchase price %") +
  scale_y_continuous("Percent") +
  scale_fill_manual(values = c("red", "green")) +
  facet_grid(RACE ~ ., scales="free") +
  theme(legend.position="bottom")

ggplot() +
  geom_bar(data = data1 %>% filter(RACE == "non-Hispanic White"),  aes(x = LOANPRC, fill = APPROVE), position = "fill")  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Loan amount/purchase price %") +
  scale_y_continuous("Percent") +
  scale_fill_manual(values = c("red", "green")) +
  facet_grid(RACE ~ ., scales="free") +
  theme(legend.position="bottom")

ggplot() +
  geom_bar(data = data1 %>% filter(MALE == "Yes"),  aes(x = LOANPRC, fill = APPROVE), position = "fill")  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Loan amount/purchase price %") +
  scale_y_continuous("Percent") +
  scale_fill_manual(values = c("red", "green")) +
  facet_grid(MALE ~ ., scales="free", labeller = as_labeller(c(Yes = "Male"))) +
  theme(legend.position="bottom")

ggplot() +
  geom_bar(data = data1 %>% filter(MALE == "No"),  aes(x = LOANPRC, fill = APPROVE), position = "fill")  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Loan amount/purchase price %") +
  scale_y_continuous("Percent") +
  scale_fill_manual(values = c("red", "green")) +
  facet_grid(MALE ~ ., scales="free", labeller = as_labeller(c(No = "Female"))) +
  theme(legend.position="bottom")


ggplot() +
  geom_bar(data = data1 %>% filter(MARRIED == "Yes"),  aes(x = LOANPRC, fill = APPROVE), position = "fill")  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Loan amount/purchase price %") +
  scale_y_continuous("Percent") +
  scale_fill_manual(values = c("red", "green")) +
  facet_grid(MARRIED ~ ., scales="free", labeller = as_labeller(c(Yes = "Married"))) +
  theme(legend.position="bottom")

ggplot() +
  geom_bar(data = data1 %>% filter(MARRIED == "No"),  aes(x = LOANPRC, fill = APPROVE), position = "fill")  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Loan amount/purchase price %") +
  scale_y_continuous("Percent") +
  scale_fill_manual(values = c("red", "green")) +
  facet_grid(MARRIED ~ ., scales="free", labeller = as_labeller(c(No = "Not Married"))) +
  theme(legend.position="bottom")
```

Descriptive statistics by Marital Status:

```{r}
data %>% select(-c(ID, BLACK, HISPAN)) %>% split(.$MARRIED) %>% map(summary)
```

Descriptive statistics by Gender:

```{r}
data %>% select(-c(ID, BLACK, HISPAN)) %>% split(.$MALE) %>% map(summary)
```


There are 3 records are missing married (MARRIED) field.

```{r}
kable(data %>% filter(MARRIED == "Unknown"))  %>%
  kable_styling(bootstrap_options = "striped")
```

There are 3 records are missing married (GDLIN) field.

```{r}
kable(data %>% filter(GDLIN == "666"))  %>%
  kable_styling(bootstrap_options = "striped")
```

There are 15 records are missing gender (MALE) field.

```{r}
kable(data %>% filter(MALE == "Unknown"))  %>%
  kable_styling(bootstrap_options = "striped")
```

```{r}
xtabs(~RACE + APPROVE, data = data)


ggplot(data,  aes(RACE)) +
  geom_bar()  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 0, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Race")
```

```{r}
xtabs(~MARRIED + APPROVE, data = data)

ggplot(data,  aes(MARRIED)) +
  geom_bar()  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 0, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Married")
```

```{r}
xtabs(~MALE + APPROVE, data = data)

ggplot(data,  aes(MALE)) +
  geom_bar()  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 0, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Male")
```

```{r}
xtabs(~MARRIED + APPROVE + RACE, data = data)


ggplot(data,  aes(RACE, fill = MALE)) +
  geom_bar()  +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 0, vjust = 1, 
                                   size = 12, hjust = 1)) + 
  xlab("Race")
```

```{r}
t1 <- theme_tufte() + theme(axis.text.x = element_text( hjust = 1)) #angle = 90,

c <- ggplot(data, aes(y = LOANPRC)) + t1 + xlab("") + ylab("Loan amount/purchase price %") + theme(legend.position="bottom")

c + geom_boxplot(aes(x=RACE, color=RACE), trim = FALSE)
c + geom_boxplot(aes(x=MARRIED, color=RACE), trim = FALSE)
c + geom_boxplot(aes(x=MALE, color=RACE), trim = FALSE)
c + geom_boxplot(aes(x=RACE, color=APPROVE), trim = FALSE)
```

```{r}
c <- ggplot(data, aes(y = OBRAT)) + t1 + xlab("") + ylab("Other obligations as a percent of total income") + theme(legend.position="bottom")
 
c + geom_boxplot(aes(x=RACE, color=RACE), trim = FALSE)
c + geom_boxplot(aes(x=MARRIED, color=RACE), trim = FALSE)
c + geom_boxplot(aes(x=MALE, color=RACE), trim = FALSE)
c + geom_boxplot(aes(x=RACE, color=APPROVE), trim = FALSE)
```

```{r}
data <- read_excel("MLD/MLD Data File.xls")
```

```{r}
#data <- read_excel("MLD Data File.xls")
```
```{r}
data <- data %>% filter(MALE != ".") %>% filter(GDLIN != "666")
data$LOANPRC <- data$LOANPRC * 100
data$GDLIN <- as.factor(data$GDLIN)

data$MARRIED <- as.factor(data$MARRIED)
data$MALE <- as.factor(data$MALE)
data$APPROVE <- as.factor(data$APPROVE)
data$BLACK <- as.factor(data$BLACK)
data$HISPAN <- as.factor(data$HISPAN)
```

## Logistic Regression Full Model

log(p/1-p) = $b_0 + b_1 * GDLIN + b_2 * OBRAT + b_3 * BLACK + b_4 * HISPAN + b_5 * MALE + b_6 * LOANPRC + b_7 * MARRIED$


```{r}
data1 <- data %>% filter(MARRIED != ".")
data1$MARRIED <- relevel(data1$MARRIED, "0")

mylogit0 <- glm(APPROVE ~  GDLIN + OBRAT + BLACK + HISPAN + MALE + LOANPRC + MARRIED , data = data1, family = "binomial")
mylogit0
```

For every one unit change in `OBRAT`, the log odds of loan approval (versus non loan approval) decreases by `r abs(mylogit0$coefficients["OBRAT"])`.

For every one unit change in `LOANPRC`, the log odds of loan approval (versus non loan approval) decreases by `r abs(mylogit0$coefficients["LOANPRC"])`.

The log odds of loan approval for applicants that meet credit guidelines increases by  `r abs(mylogit0$coefficients["GDLIN1"])`.

The log odds of loan approval for married applicants increases by  `r abs(mylogit0$coefficients["MARRIED1"])`.

The log odds of loan approval for Black applicants decreases by  `r abs(mylogit0$coefficients["BLACK1"])`.

The log odds of loan approval for Hispanic applicants decreases by  `r abs(mylogit0$coefficients["HISPAN1"])`.

### Statistical Tests for Individual Predictors

```{r}
varImp(mylogit0)

regTermTest(mylogit0, "GDLIN")
regTermTest(mylogit0, "OBRAT")
regTermTest(mylogit0, "BLACK")
regTermTest(mylogit0, "HISPAN")
regTermTest(mylogit0, "MALE")
regTermTest(mylogit0, "LOANPRC")
regTermTest(mylogit0, "MARRIED")
pR2(mylogit0)  # look for 'McFadden'
```

## Logistic Regression Reduced Model 1

```{r}
mylogit1 <- glm(APPROVE ~  GDLIN + OBRAT + BLACK + HISPAN + MARRIED + LOANPRC , data = data1, family = "binomial")
mylogit1
regTermTest(mylogit1, "GDLIN")
regTermTest(mylogit1, "OBRAT")
regTermTest(mylogit1, "BLACK")
regTermTest(mylogit1, "HISPAN")
regTermTest(mylogit1, "MARRIED")
regTermTest(mylogit1, "LOANPRC")
```

For every one unit change in `OBRAT`, the log odds of loan approval (versus non loan approval) decreases by `r abs(mylogit1$coefficients["OBRAT"])`.

For every one unit change in `LOANPRC`, the log odds of loan approval (versus non loan approval) decreases by `r abs(mylogit1$coefficients["LOANPRC"])`.

The log odds of loan approval for applicants that meet credit guidelines increase by `r abs(mylogit1$coefficients["GDLIN1"])`.

The log odds of loan approval for married applicants decreases by  `r abs(mylogit0$coefficients["MARRIED1"])`.

The log odds of loan approval for non married applicants decreases by  `r abs(mylogit0$coefficients["MARRIED0"])`.

The log odds of loan approval for Black applicants decreases by `r abs(mylogit1$coefficients["BLACK1"])`.

The log odds of loan approval for Hispanic applicants decreases by  `r abs(mylogit1$coefficients["HISPAN1"])`.

### CIs using profiled log-likelihood

```{r}
confint(mylogit1)
```

### CIs using standard errors

```{r}
confint.default(mylogit1)
```

### Odds ratios only

```{r}
exp(coef(mylogit1))
```

### Odds ratios and 95% CI

```{r}
exp(cbind(OR = coef(mylogit1), confint(mylogit1)))
```

### Model comparison

```{r}
anova(mylogit0, mylogit1, test ="Chisq")
lrtest(mylogit0, mylogit1)
```

## Logistic Regression Reduced Model 2

```{r}
mylogit2 <- glm(APPROVE ~  GDLIN + OBRAT + BLACK + HISPAN  + LOANPRC , data = data1, family = "binomial")
mylogit2
regTermTest(mylogit1, "GDLIN")
regTermTest(mylogit1, "OBRAT")
regTermTest(mylogit1, "BLACK")
regTermTest(mylogit1, "HISPAN")
regTermTest(mylogit1, "LOANPRC")
```

For every one unit change in `OBRAT`, the log odds of loan approval (versus non loan approval) decreases by `r abs(mylogit2$coefficients["OBRAT"])`.

For every one unit change in `LOANPRC`, the log odds of loan approval (versus non loan approval) decreases by `r abs(mylogit2$coefficients["LOANPRC"])`.

The log odds of loan approval for applicants that meet credit guidelines increases by  `r abs(mylogit2$coefficients["GDLIN1"])`.

The log odds of loan approval for Black applicants decreases by  `r abs(mylogit2$coefficients["BLACK1"])`.

The log odds of loan approval for Hispanic applicants decreases by  `r abs(mylogit2$coefficients["HISPAN1"])`.

### CIs using profiled log-likelihood

```{r}
confint(mylogit2)
```

### CIs using standard errors

```{r}
confint.default(mylogit2)
```

### Odds ratios only

```{r}
exp(coef(mylogit2))
```

### Odds ratios and 95% CI

```{r}
exp(cbind(OR = coef(mylogit2), confint(mylogit2)))
```

### Model comparison

```{r}
anova(mylogit0, mylogit2, test ="Chisq")
lrtest(mylogit0, mylogit2)
```

```{r}
AVG_LOANPRC <- mean(data$LOANPRC)
AVG_OBRAT <- mean(data$OBRAT)

#newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 96)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(1), HISPAN = factor(0), MALE = factor(rep(0:1, each = 96)), LOANPRC = 86.05))

newdata21 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 192)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(1), HISPAN = factor(0), MARRIED = factor(rep(0:1, each = 96)), LOANPRC = AVG_LOANPRC))
newdata22 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 192)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(1), MARRIED = factor(rep(0:1, each = 96)), LOANPRC = AVG_LOANPRC))
newdata23 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 192)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(0), MARRIED = factor(rep(0:1, each = 96)), LOANPRC = AVG_LOANPRC))


newdata24 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 200)), LOANPRC = rep(seq(from = 2.105, to = 257.143, length.out = 100)), BLACK = factor(1), HISPAN = factor(0), MARRIED = factor(rep(0:1, each = 100)), OBRAT = AVG_OBRAT))
newdata25 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 200)), LOANPRC = rep(seq(from = 2.105, to = 257.143, length.out = 100)), BLACK = factor(0), HISPAN = factor(1), MARRIED = factor(rep(0:1, each = 100)), OBRAT = AVG_OBRAT))
newdata26 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 200)), LOANPRC = rep(seq(from = 2.105, to = 257.143, length.out = 100)), BLACK = factor(0), HISPAN = factor(0), MARRIED = factor(rep(0:1, each = 100)), OBRAT = AVG_OBRAT))


newdata31 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 96)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(1), HISPAN = factor(0), LOANPRC = AVG_LOANPRC))
newdata32 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 96)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(1), LOANPRC = AVG_LOANPRC))
newdata33 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 96)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(0), LOANPRC = AVG_LOANPRC))


newdata34 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 100)), LOANPRC = rep(seq(from = 2.105, to = 257.143, length.out = 100)), BLACK = factor(1), HISPAN = factor(0), OBRAT = AVG_OBRAT))
newdata35 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 100)), LOANPRC = rep(seq(from = 2.105, to = 257.143, length.out = 100)), BLACK = factor(0), HISPAN = factor(1), OBRAT = AVG_OBRAT))
newdata36 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 100)), LOANPRC = rep(seq(from = 2.105, to = 257.143, length.out = 100)), BLACK = factor(0), HISPAN = factor(0), OBRAT = AVG_OBRAT))
```

```{r eval=FALSE}     
# non-Hispanic Black Male
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 96)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(1), HISPAN = factor(0), MALE = factor(1), LOANPRC = AVG_LOANPRC))
newdata3 <- cbind(newdata2, predict(mylogit1, newdata = newdata2, type = "link", se = TRUE))
newdata3 <- within(newdata3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

ggplot(newdata3, aes(x = OBRAT, y = PredictedProb)) + 
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = GDLIN), alpha = 0.2) +
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("non-Hispanic Black Male Predicted probabilities")


# non-Hispanic Black Female
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 96)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(1), HISPAN = factor(0), MALE = factor(0), LOANPRC = AVG_LOANPRC))
newdata3 <- cbind(newdata2, predict(mylogit1, newdata = newdata2, type = "link", se = TRUE))
newdata3 <- within(newdata3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

ggplot(newdata3, aes(x = OBRAT, y = PredictedProb)) + 
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = GDLIN), alpha = 0.2) +
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("non-Hispanic Black Female Predicted probabilities")
```

```{r eval=FALSE}     
# Hispanic Male
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 96)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(1), MALE = factor(1), LOANPRC = AVG_LOANPRC))
newdata3 <- cbind(newdata2, predict(mylogit1, newdata = newdata2, type = "link", se = TRUE))
newdata3 <- within(newdata3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

ggplot(newdata3, aes(x = OBRAT, y = PredictedProb)) + 
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = GDLIN), alpha = 0.2) +
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("Hispanic Male Predicted probabilities")
```

```{r eval=FALSE}     
# Hispanic Female
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 96)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(1), MALE = factor(0), LOANPRC = AVG_LOANPRC))
newdata3 <- cbind(newdata2, predict(mylogit1, newdata = newdata2, type = "link", se = TRUE))
newdata3 <- within(newdata3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

ggplot(newdata3, aes(x = OBRAT, y = PredictedProb)) + 
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = GDLIN), alpha = 0.2) +
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("Hispanic Female Predicted probabilities")
```

```{r eval=FALSE}     
# White Male
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 96)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(0), MALE = factor(1), LOANPRC = AVG_LOANPRC))
newdata3 <- cbind(newdata2, predict(mylogit1, newdata = newdata2, type = "link", se = TRUE))
newdata3 <- within(newdata3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

ggplot(newdata3, aes(x = OBRAT, y = PredictedProb)) + geom_ribbon(aes(ymin = LL,
                                                                      ymax = UL, fill = GDLIN), alpha = 0.2) + geom_line(aes(colour = GDLIN),
                                                                                                                         size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("non-Hispanic White Male Predicted probabilities")
```

```{r eval=FALSE}     
# White Female
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 96)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(0), MALE = factor(0), LOANPRC = AVG_LOANPRC))
newdata3 <- cbind(newdata2, predict(mylogit1, newdata = newdata2, type = "link", se = TRUE))
newdata3 <- within(newdata3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

ggplot(newdata3, aes(x = OBRAT, y = PredictedProb)) + 
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = GDLIN), alpha = 0.2) +
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("non-Hispanic White Female Predicted probabilities")
```


```{r eval=FALSE}     
# non-Hispanic Black Male
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 100)), LOANPRC = rep(seq(from = 2.105, to = 257.143, length.out = 100)), BLACK = factor(1), HISPAN = factor(0), MALE = factor(1), OBRAT = AVG_OBRAT))
newdata3 <- cbind(newdata2, predict(mylogit1, newdata = newdata2, type = "link", se = TRUE))
newdata3 <- within(newdata3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

ggplot(newdata3, aes(x = LOANPRC, y = PredictedProb)) + 
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = GDLIN), alpha = 0.2) +
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("non-Hispanic Black Male Predicted probabilities")


# non-Hispanic Black Female
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 100)), LOANPRC = rep(seq(from = 2.105, to = 257.143, length.out = 100)), BLACK = factor(1), HISPAN = factor(0), MALE = factor(0), OBRAT = AVG_OBRAT))
newdata3 <- cbind(newdata2, predict(mylogit1, newdata = newdata2, type = "link", se = TRUE))
newdata3 <- within(newdata3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

ggplot(newdata3, aes(x = LOANPRC, y = PredictedProb)) + 
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = GDLIN), alpha = 0.2) +
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("non-Hispanic Black Female Predicted probabilities")
```

```{r eval=FALSE}     
# Hispanic Male
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 100)), LOANPRC = rep(seq(from = 2.105, to = 257.143, length.out = 100)), BLACK = factor(0), HISPAN = factor(1), MALE = factor(1), OBRAT = AVG_OBRAT))
newdata3 <- cbind(newdata2, predict(mylogit1, newdata = newdata2, type = "link", se = TRUE))
newdata3 <- within(newdata3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

ggplot(newdata3, aes(x = LOANPRC, y = PredictedProb)) + 
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = GDLIN), alpha = 0.2) +
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("Hispanic Male Predicted probabilities")
```

```{r eval=FALSE}     
# Hispanic Female
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 100)), LOANPRC = rep(seq(from = 2.105, to = 257.143, length.out = 100)), BLACK = factor(0), HISPAN = factor(1), MALE = factor(1), OBRAT = AVG_OBRAT))
newdata3 <- cbind(newdata2, predict(mylogit1, newdata = newdata2, type = "link", se = TRUE))
newdata3 <- within(newdata3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

ggplot(newdata3, aes(x = LOANPRC, y = PredictedProb)) + 
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = GDLIN), alpha = 0.2) +
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("Hispanic Female Predicted probabilities")
```

```{r eval=FALSE}     
# White Male
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 100)), LOANPRC = rep(seq(from = 2.105, to = 257.143, length.out = 100)), BLACK = factor(0), HISPAN = factor(0), MALE = factor(1), OBRAT = AVG_OBRAT))
newdata3 <- cbind(newdata2, predict(mylogit1, newdata = newdata2, type = "link", se = TRUE))
newdata3 <- within(newdata3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

ggplot(newdata3, aes(x = LOANPRC, y = PredictedProb)) +
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = GDLIN), alpha = 0.2) +
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("non-Hispanic White Male Predicted probabilities")
```

```{r eval=FALSE}     
# White Female
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 100)), LOANPRC = rep(seq(from = 2.105, to = 257.143, length.out = 100)), BLACK = factor(0), HISPAN = factor(0), MALE = factor(0), OBRAT = AVG_OBRAT))
newdata3 <- cbind(newdata2, predict(mylogit1, newdata = newdata2, type = "link", se = TRUE))
newdata3 <- within(newdata3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

ggplot(newdata3, aes(x = LOANPRC, y = PredictedProb)) + 
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = GDLIN), alpha = 0.2) +
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("non-Hispanic White Female Predicted probabilities")
```

## Probit Regression Full Model

```{r}
myprobit0 <- glm(APPROVE ~  GDLIN + OBRAT + BLACK + HISPAN + MALE + LOANPRC + MARRIED, data = data1, family = binomial(link = "probit"))
myprobit0
```

### Statistical Tests for Individual Predictors

```{r}
regTermTest(myprobit0, "GDLIN")
regTermTest(myprobit0, "OBRAT")
regTermTest(myprobit0, "BLACK")
regTermTest(myprobit0, "HISPAN")
regTermTest(myprobit0, "MALE")
regTermTest(myprobit0, "LOANPRC")
regTermTest(myprobit0, "MARRIED")
```

## Probit Regression Reduced Model 1

```{r}
myprobit1 <- glm(APPROVE ~  GDLIN + OBRAT + BLACK + HISPAN + MARRIED + LOANPRC , data = data1, family = binomial(link = "probit"))
myprobit1
varImp(myprobit1)
pR2(myprobit1)  # look for 'McFadden'
```

```{r}
anova(myprobit0, myprobit1, test ="Chisq")
lrtest(myprobit0, myprobit1)
```

## Probit Regression Reduced Model 2

```{r}
myprobit2 <- glm(APPROVE ~  GDLIN + OBRAT + BLACK + HISPAN + LOANPRC, data = data1, family = binomial(link = "probit"))
myprobit2
varImp(myprobit2)
pR2(myprobit2)  # look for 'McFadden'
```

```{r}
anova(myprobit0, myprobit2, test ="Chisq")
lrtest(myprobit0, myprobit2)
```

<!-- ## Probit Regression Reduced Model 3 -->

<!-- ```{r} -->
<!-- myprobit3 <- glm(APPROVE ~  GDLIN  + BLACK + HISPAN + MALE + LOANPRC , data = data, family = binomial(link = "probit")) -->
<!-- myprobit3 -->
<!-- varImp(myprobit3) -->
<!-- pR2(myprobit3)  # look for 'McFadden' -->
<!-- ``` -->

<!-- ```{r} -->
<!-- anova(myprobit0, myprobit3, test ="Chisq") -->
<!-- lrtest(myprobit0, myprobit3) -->
<!-- ``` -->

<!-- ## Probit Regression Reduced Model 4 -->

<!-- ```{r} -->
<!-- myprobit4 <- glm(APPROVE ~  GDLIN  + BLACK + HISPAN  + LOANPRC , data = data, family = binomial(link = "probit")) -->
<!-- myprobit4 -->
<!-- varImp(myprobit4) -->
<!-- pR2(myprobit4)  # look for 'McFadden' -->
<!-- ``` -->

<!-- ```{r} -->
<!-- anova(myprobit1, myprobit4, test ="Chisq") -->
<!-- lrtest(myprobit1, myprobit4) -->
<!-- ``` -->


```{r eval=FALSE}
# non-Hispanic Black Male/Female
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 192)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(1), HISPAN = factor(0), MALE = factor(rep(0:1, each = 96)), LOANPRC = AVG_LOANPRC))
newdata3 <- cbind(newdata2, predict(myprobit1, newdata2, type = "response", se.fit = TRUE)[-3])
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")
newdata3$MALE <- ifelse(newdata3$MALE == 1, "Male", "Female")

ggplot(newdata3, aes(x = OBRAT, y = fit)) + 
  geom_line(aes(colour = MALE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("non-Hispanic Black Predicted probabilities") + 
  facet_grid(GDLIN ~ .) +
  ylab("PredictedProb")


ggplot(newdata3, aes(x = OBRAT, y = fit)) + 
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("non-Hispanic Black Predicted probabilities") + 
  facet_wrap(MALE ~ .) +
  ylab("PredictedProb")
```

```{r eval=FALSE}
# Hispanic Male/Female
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 192)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(1), MALE = factor(rep(0:1, each = 96)), LOANPRC = AVG_LOANPRC))
newdata3 <- cbind(newdata2, predict(myprobit1, newdata2, type = "response", se.fit = TRUE)[-3])
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")
newdata3$MALE <- ifelse(newdata3$MALE == 1, "Male", "Female")

ggplot(newdata3, aes(x = OBRAT, y = fit)) + 
  geom_line(aes(colour = MALE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("Hispanic Predicted probabilities") + 
  facet_grid(GDLIN ~ .) +
  ylab("PredictedProb")


ggplot(newdata3, aes(x = OBRAT, y = fit)) + 
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("Hispanic Predicted probabilities") + 
  facet_wrap(MALE ~ .) +
  ylab("PredictedProb")
```

```{r eval=FALSE}
# White Male/Female
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 192)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(0), MALE = factor(rep(0:1, each = 96)), LOANPRC = AVG_LOANPRC))
newdata3 <- cbind(newdata2, predict(myprobit1, newdata2, type = "response", se.fit = TRUE)[-3])
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")
newdata3$MALE <- ifelse(newdata3$MALE == 1, "Male", "Female")

ggplot(newdata3, aes(x = OBRAT, y = fit)) + 
  geom_line(aes(colour = MALE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("non-Hispanic White Predicted probabilities") + 
  facet_grid(GDLIN ~ .) +
  ylab("PredictedProb")


ggplot(newdata3, aes(x = OBRAT, y = fit)) + 
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("non-Hispanic White Predicted probabilities") + 
  facet_wrap(MALE ~ .) +
  ylab("PredictedProb")
```

```{r eval=FALSE}
# non-Hispanic Black Male/Female
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 200)), LOANPRC = rep(seq(from = 2.105, to = 257.143, length.out = 100)), BLACK = factor(1), HISPAN = factor(0), MALE = factor(rep(0:1, each = 100)), OBRAT = AVG_OBRAT))
newdata3 <- cbind(newdata2, predict(myprobit1, newdata2, type = "response", se.fit = TRUE)[-3])
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")
newdata3$MALE <- ifelse(newdata3$MALE == 1, "Male", "Female")

ggplot(newdata3, aes(x = LOANPRC, y = fit)) + 
  geom_line(aes(colour = MALE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("non-Hispanic Black Predicted probabilities") + 
  facet_grid(GDLIN ~ .) +
  ylab("PredictedProb")


ggplot(newdata3, aes(x = LOANPRC, y = fit)) + 
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("non-Hispanic Black Predicted probabilities") + 
  facet_wrap(MALE ~ .) +
  ylab("PredictedProb")
```

```{r eval=FALSE}
# non-Hispanic Hispanic Male/Female
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 200)), LOANPRC = rep(seq(from = 2.105, to = 257.143, length.out = 100)), BLACK = factor(0), HISPAN = factor(1), MALE = factor(rep(0:1, each = 100)), OBRAT = AVG_OBRAT))
newdata3 <- cbind(newdata2, predict(myprobit1, newdata2, type = "response", se.fit = TRUE)[-3])
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")
newdata3$MALE <- ifelse(newdata3$MALE == 1, "Male", "Female")

ggplot(newdata3, aes(x = LOANPRC, y = fit)) + 
  geom_line(aes(colour = MALE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("Hispanic Predicted probabilities") + 
  facet_grid(GDLIN ~ .) +
  ylab("PredictedProb")


ggplot(newdata3, aes(x = LOANPRC, y = fit)) + 
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("Hispanic Predicted probabilities") + 
  facet_wrap(MALE ~ .) +
  ylab("PredictedProb")
```

```{r eval=FALSE}
# White Male/Female
newdata2 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 200)), LOANPRC = rep(seq(from = 2.105, to = 257.143, length.out = 100)), BLACK = factor(0), HISPAN = factor(0), MALE = factor(rep(0:1, each = 100)), OBRAT = AVG_OBRAT))
newdata3 <- cbind(newdata2, predict(myprobit1, newdata2, type = "response", se.fit = TRUE)[-3])
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")
newdata3$MALE <- ifelse(newdata3$MALE == 1, "Male", "Female")

ggplot(newdata3, aes(x = LOANPRC, y = fit)) + 
  geom_line(aes(colour = MALE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("non-Hispanic White Black Predicted probabilities") + 
  facet_grid(GDLIN ~ .) +
  ylab("PredictedProb")


ggplot(newdata3, aes(x = LOANPRC, y = fit)) + 
  geom_line(aes(colour = GDLIN), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("non-Hispanic White Black Predicted probabilities") + 
  facet_wrap(MALE ~ .) +
  ylab("PredictedProb")
```

## TBD

```{r}
#
# newdata21 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 192)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(1), HISPAN = factor(0), MALE = factor(rep(0:1, each = 96)), LOANPRC = AVG_LOANPRC))
# newdata22 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 192)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(1), MALE = factor(rep(0:1, each = 96)), LOANPRC = AVG_LOANPRC))
# newdata23 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 192)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(0), MALE = factor(rep(0:1, each = 96)), LOANPRC = AVG_LOANPRC))

newdata2 <- rbind(newdata21, newdata22, newdata23)
newdata3 <- cbind(newdata2, predict(mylogit1, newdata = newdata2, type = "link", se = TRUE))
newdata3 <- within(newdata3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})
#newdata3 <- cbind(newdata2, predict(mylogit, newdata = newdata2, type = "response", se = TRUE))


newdata3$RACE <- ifelse(newdata3$BLACK == 1, "non-Hispanic Black", ifelse(newdata3$HISPAN == 1, "Hispanic", "non-Hispanic White"))
newdata3$RACE  <- as.factor(newdata3$RACE)
#newdata3$MALE <- ifelse(newdata3$MALE == 1, "Male", "Female")
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")
newdata3$MARRIED <- ifelse(newdata3$MARRIED == 1, "Married", "Not Married")


ggplot(newdata3, aes(x = OBRAT, y = PredictedProb)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("Predicted probabilities") + 
  facet_wrap(GDLIN ~ MARRIED, scales="free") +
  ylab("PredictedProb")
```


```{r}
newdata2 <- rbind(newdata24, newdata25, newdata26)
newdata3 <- cbind(newdata2, predict(mylogit1, newdata = newdata2, type = "link", se = TRUE))
newdata3 <- within(newdata3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

newdata3$RACE <- ifelse(newdata3$BLACK == 1, "non-Hispanic Black", ifelse(newdata3$HISPAN == 1, "Hispanic", "non-Hispanic White"))
newdata3$RACE  <- as.factor(newdata3$RACE)
#newdata3$MALE <- ifelse(newdata3$MALE == 1, "Male", "Female")
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")
newdata3$MARRIED <- ifelse(newdata3$MARRIED == 1, "Married", "Not Married")


ggplot(newdata3, aes(x = LOANPRC, y = PredictedProb)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("Predicted probabilities") + 
  facet_wrap(GDLIN ~ MARRIED, scales="free") +
  ylab("PredictedProb")
```



```{r}
newdata2 <- rbind(newdata31, newdata32, newdata33)
newdata3 <- cbind(newdata2, predict(mylogit2, newdata = newdata2, type = "link", se = TRUE))
newdata3 <- within(newdata3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

newdata3$RACE <- ifelse(newdata3$BLACK == 1, "non-Hispanic Black", ifelse(newdata3$HISPAN == 1, "Hispanic", "non-Hispanic White"))
newdata3$RACE  <- as.factor(newdata3$RACE)
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")

ggplot(newdata3, aes(x = OBRAT, y = PredictedProb)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("Predicted probabilities") + 
  facet_wrap(GDLIN ~ RACE) +
  ylab("PredictedProb")

ggplot(newdata3, aes(x = OBRAT, y = PredictedProb)) + 
  geom_line(aes(colour = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("Predicted probabilities") + 
  facet_wrap(GDLIN ~ ., scales="free") +
  ylab("PredictedProb")
```


```{r}
newdata2 <- rbind(newdata34, newdata35, newdata36)
newdata3 <- cbind(newdata2, predict(mylogit2, newdata = newdata2, type = "link", se = TRUE))
newdata3 <- within(newdata3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

newdata3$RACE <- ifelse(newdata3$BLACK == 1, "non-Hispanic Black", ifelse(newdata3$HISPAN == 1, "Hispanic", "non-Hispanic White"))
newdata3$RACE  <- as.factor(newdata3$RACE)
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")

ggplot(newdata3, aes(x = LOANPRC, y = PredictedProb)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("Predicted probabilities") + 
  facet_wrap(GDLIN ~ RACE) +
  ylab("PredictedProb")

ggplot(newdata3, aes(x = LOANPRC, y = PredictedProb)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("Predicted probabilities") + 
  facet_wrap(GDLIN ~ ., scales="free") +
  ylab("PredictedProb")
```


## TBD

```{r}
newdata2 <- rbind(newdata21, newdata22, newdata23)
newdata3 <- cbind(newdata2, predict(myprobit1, newdata2, type = "response", se.fit = TRUE)[-3])

newdata3$RACE <- ifelse(newdata3$BLACK == 1, "non-Hispanic Black", ifelse(newdata3$HISPAN == 1, "Hispanic", "non-Hispanic White"))
newdata3$RACE  <- as.factor(newdata3$RACE)
#newdata3$MALE <- ifelse(newdata3$MALE == 1, "Male", "Female")
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")
newdata3$MARRIED <- ifelse(newdata3$MARRIED == 1, "Married", "Not Married")

ggplot(newdata3, aes(x = OBRAT, y = fit)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("Predicted probabilities") + 
  facet_wrap(GDLIN ~ MARRIED, scales="free") +
  ylab("PredictedProb")
```


```{r}
newdata2 <- rbind(newdata24, newdata25, newdata26)
newdata3 <- cbind(newdata2, predict(myprobit1, newdata2, type = "response", se.fit = TRUE)[-3])

newdata3$RACE <- ifelse(newdata3$BLACK == 1, "non-Hispanic Black", ifelse(newdata3$HISPAN == 1, "Hispanic", "non-Hispanic White"))
newdata3$RACE  <- as.factor(newdata3$RACE)
#newdata3$MALE <- ifelse(newdata3$MALE == 1, "Male", "Female")
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")
newdata3$MARRIED <- ifelse(newdata3$MARRIED == 1, "Married", "Not Married")


ggplot(newdata3, aes(x = LOANPRC, y = fit)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("Predicted probabilities") + 
  facet_wrap(GDLIN ~ MARRIED, scales="free") +
  ylab("PredictedProb")
```


```{r}
newdata2 <- rbind(newdata31, newdata32, newdata33)
newdata3 <- cbind(newdata2, predict(myprobit2, newdata2, type = "response", se.fit = TRUE)[-3])

newdata3$RACE <- ifelse(newdata3$BLACK == 1, "non-Hispanic Black", ifelse(newdata3$HISPAN == 1, "Hispanic", "non-Hispanic White"))
newdata3$RACE  <- as.factor(newdata3$RACE)
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")

ggplot(newdata3, aes(x = OBRAT, y = fit)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("Predicted probabilities") + 
  facet_wrap(GDLIN ~ RACE) +
  ylab("PredictedProb")

ggplot(newdata3, aes(x = OBRAT, y = fit)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("Predicted probabilities") + 
  facet_wrap(GDLIN ~ ., scales="free") +
  ylab("PredictedProb")
```


```{r}
newdata2 <- rbind(newdata34, newdata35, newdata36)
newdata3 <- cbind(newdata2, predict(myprobit2, newdata2, type = "response", se.fit = TRUE)[-3])

newdata3$RACE <- ifelse(newdata3$BLACK == 1, "non-Hispanic Black", ifelse(newdata3$HISPAN == 1, "Hispanic", "non-Hispanic White"))
newdata3$RACE  <- as.factor(newdata3$RACE)
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")

ggplot(newdata3, aes(x = LOANPRC, y = fit)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("Predicted probabilities") + 
  facet_wrap(GDLIN ~ RACE) +
  ylab("PredictedProb")

ggplot(newdata3, aes(x = LOANPRC, y = fit)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("Predicted probabilities") + 
  facet_wrap(GDLIN ~ ., scales="free") +
  ylab("PredictedProb")
```