---
title: "Lending Mortage Analysis - Modeling"
author: "Marjorie Blanco, Joe Thompson, Haodi Tu"
output:
  pdf_document: default
  html_document: default
subtitle: Exploratory Data Analysis
geometry: left=2cm,right=3cm,top=2cm,bottom=2cm
---

```{r warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(scipen=999)
library(readr)
library(ggplot2)
library(readxl)
library(kableExtra)
library(dplyr)
library(tidyverse)
library(purrr)
library(ggthemes)
library(aod)
library(lmtest)
library(survey)
library(caret)
library(pscl)
library(survey)
library(stargazer)
library(foreign)
```


```{r}
data <- read_excel("MLD/MLD Data File.xls")
data <- data.frame(data)

data <- data %>% filter(MALE != ".") %>% filter(GDLIN != "666") %>% filter(LOANPRC < 1)
data$LOANPRC <- data$LOANPRC * 100

stargazer(data, type = "text", title="Descriptive statistics", digits=2)

stargazer(data %>% filter(BLACK == 0) %>% filter(HISPAN == 0) %>% select(-c(BLACK, HISPAN)), type = "text", title="Descriptive statistics - non-Hispanic White", digits=2)

stargazer(data %>% filter(BLACK == 1) %>% filter(HISPAN == 0) %>% select(-BLACK), type = "text", title="Descriptive statistics - non-Hispanic Black", digits=2)

stargazer(data %>% filter(BLACK == 0) %>% filter(HISPAN == 1) %>% select(-HISPAN), type = "text", title="Descriptive statistics - Hispanic", digits=2)

```

```{r}
#data <- read_excel("MLD Data File.xls")
```

```{r}
#http://people.umass.edu/biep640w/pdf/R-for-Logistic-Regression.pdf
#https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf
#https://www.r-bloggers.com/evaluating-logistic-regression-models/
```

```{r}
data$GDLIN <- as.factor(data$GDLIN)
data$MARRIED <- as.factor(data$MARRIED)
data$MALE <- as.factor(data$MALE)
data$APPROVE <- as.factor(data$APPROVE)
data$BLACK <- as.factor(data$BLACK)
data$HISPAN <- as.factor(data$HISPAN)
```

```{r}

```


```{r eval=FALSE}
#Descriptive statistics by Race:




```

```{r}
### Overall
summary(data)

data$RACE <- ifelse(data$BLACK == 1, "non-Hispanic Black", ifelse(data$HISPAN == 1, "Hispanic", "non-Hispanic White"))
data$RACE <-as.factor(data$RACE)
data %>% select(-c(BLACK, HISPAN)) %>% split(.$RACE) %>% map(summary) 

```

```{r}
AVG_LOANPRC <- mean(data$LOANPRC)
AVG_OBRAT <- mean(data$OBRAT)

newdata21 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 192)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(1), HISPAN = factor(0), MARRIED = factor(rep(0:1, each = 96)), LOANPRC = AVG_LOANPRC))
newdata22 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 192)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(1), MARRIED = factor(rep(0:1, each = 96)), LOANPRC = AVG_LOANPRC))
newdata23 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 192)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(0), MARRIED = factor(rep(0:1, each = 96)), LOANPRC = AVG_LOANPRC))


newdata24 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 200)), OBRAT = AVG_OBRAT, BLACK = factor(1), HISPAN = factor(0), MARRIED = factor(rep(0:1, each = 100)), LOANPRC = rep(seq(from = 2.105, to = 100, length.out = 100))))
newdata25 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 200)), OBRAT = AVG_OBRAT, BLACK = factor(0), HISPAN = factor(1), MARRIED = factor(rep(0:1, each = 100)), LOANPRC = rep(seq(from = 2.105, to = 100, length.out = 100))))
newdata26 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 200)), OBRAT = AVG_OBRAT, BLACK = factor(0), HISPAN = factor(0), MARRIED = factor(rep(0:1, each = 100)), LOANPRC = rep(seq(from = 2.105, to = 100, length.out = 100))))


newdata31 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 96)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(1), HISPAN = factor(0), LOANPRC = AVG_LOANPRC))
newdata32 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 96)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(1), LOANPRC = AVG_LOANPRC))
newdata33 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 96)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(0), LOANPRC = AVG_LOANPRC))


newdata34 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 100)), OBRAT = AVG_OBRAT, BLACK = factor(1), HISPAN = factor(0), LOANPRC = rep(seq(from = 2.105, to = 100, length.out = 100))))
newdata35 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 100)), OBRAT = AVG_OBRAT, BLACK = factor(0), HISPAN = factor(1), LOANPRC = rep(seq(from = 2.105, to = 100, length.out = 100))))
newdata36 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 100)), OBRAT = AVG_OBRAT, BLACK = factor(0), HISPAN = factor(0), LOANPRC = rep(seq(from = 2.105, to = 100, length.out = 100))))


newdata41 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 1)), OBRAT = AVG_OBRAT, BLACK = factor(1), HISPAN = factor(0), LOANPRC = AVG_LOANPRC))
newdata42 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 1)), OBRAT = AVG_OBRAT, BLACK = factor(0), HISPAN = factor(1), LOANPRC = AVG_LOANPRC))
newdata43 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 1)), OBRAT = AVG_OBRAT, BLACK = factor(0), HISPAN = factor(0), LOANPRC = AVG_LOANPRC))
```

## Data Discription

The data set includes the following variables: 

+ APPROVE = 1 if mortgage loan was approved, = 0 otherwise
+ GDLIN = 1 if credit history meets guidelines, = 0 otherwise
+ LOANPRC = loan amount/purchase price
+ OBRAT = other obligations as a percent of total income
+ MALE = 1 if male, = 0 otherwise
+ MARRIED = 1 if married, = 0 otherwise
+ BLACK = 1 if black, = 0 otherwise
+ HISPAN = 1 if Hispanic, = 0 otherwise

## Logistic Model

$$log(p/1-p) = \beta_0 + \beta_1 * GDLIN + \beta_2 * OBRAT + \beta_3 * BLACK + \beta_4 * HISPAN + \beta_5 * LOANPRC$$


```{r}
library(nnet)
LogitModel <- glm(APPROVE ~  GDLIN + OBRAT + BLACK + HISPAN  + LOANPRC , data = data, family = "binomial")

data1  <- data %>% mutate(APPROVE = ifelse(APPROVE == 1, "Yes", ifelse(APPROVE == 0, "No", "Unknown")))
data1$APPROVE <- as.factor(data1$APPROVE)
data1$APPROVE = relevel(data1$APPROVE, ref = "Yes")
multi1 = multinom(APPROVE ~  GDLIN + OBRAT + BLACK + HISPAN  + LOANPRC , data = data1)

#Generate Log-Likelihood
logLik(LogitModel)

stargazer(LogitModel,
          type = "text",
          digits = 3,
          header = T,
          title = "Logistic Regression")

OR <- function(x) exp(x)

stargazer(LogitModel, 
          type="text", 
          apply.coef = OR, 
          title = "Odds Ratio")

stargazer(multi1, type="text")
```

```{r}
or <- exp(coef(LogitModel))
```


Controlling for other variables, applicants who meet the credit guidelines have an estimate of `r or[2]` times higher odds of loan approval than the applicants who do not meet the credit guidelines.  We find that after controlling for all other variables, odds of loan approval for White applicants are 2.46 times and 3.24 times greater in comparison to Blacks and Hispanics respectively.


For every one unit change in `OBRAT`, the log odds of loan approval (versus non loan approval) decreases by `r abs(LogitModel$coefficients["OBRAT"])`.

For every one unit change in `LOANPRC`, the log odds of loan approval (versus non loan approval) decreases by `r abs(LogitModel$coefficients["LOANPRC"])`.

The log odds of loan approval for applicants that meet credit guidelines increases by  `r abs(LogitModel$coefficients["GDLIN1"])`.

The log odds of loan approval for Black applicants decreases by  `r abs(LogitModel$coefficients["BLACK1"])`.

The log odds of loan approval for Hispanic applicants decreases by  `r abs(LogitModel$coefficients["HISPAN1"])`.

For example, for a black person whose credit history meets guidline (GDLIN = 1), loan amount price is 100 (LOANPRC = 100) and other obligations as a percent of total income is none (OBRAT = 0), the log odds of loan approval is `r (round(1 / (1 + (exp(-(LogitModel$coefficients["(Intercept)"] + LogitModel$coefficients["GDLIN1"] + LogitModel$coefficients["BLACK1"] + 100 * LogitModel$coefficients["LOANPRC"])))), 4))*100`%

### CIs using profiled log-likelihood

```{r}
confint(LogitModel)
```

### CIs using standard errors

```{r}
confint.default(LogitModel)
```

### Odds ratios only

```{r}
exp(coef(LogitModel))
```

### Odds ratios and 95% CI

```{r}
exp(cbind(OR = coef(LogitModel), confint(LogitModel)))
```

```{r}
newdata2 <- rbind(newdata31, newdata32, newdata33)
newdata3 <- cbind(newdata2, predict(LogitModel, newdata = newdata2, type = "link", se = TRUE)[1])
newdata3$PredictedProb <- plogis(newdata3$fit)
  # LL <- plogis(fit - (1.96 * se.fit))
  # UL <- plogis(fit + (1.96 * se.fit))
#})

newdata4 <- rbind(newdata41, newdata42, newdata43)
newdata5 <- cbind(newdata4, predict(LogitModel, newdata = newdata4, type = "link", se = TRUE)[1])
newdata5$PredictedProb <- plogis(newdata5$fit)

# newdata5 <- within(newdata5, {
#   PredictedProb <- plogis(fit)
  # LL <- plogis(fit - (1.96 * se.fit))
  # UL <- plogis(fit + (1.96 * se.fit))
# })

#predict(LogitModel, newdata = newdata5, type = "response", se = TRUE)
```

```{r eval=FALSE}
sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 1, 0, 75.44245)))
sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 1, 0, 75.44245)))
sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 1, 75.44245)))
sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 1, 75.44245)))
sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 0, 75.44245)))
sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 0, 75.44245)))


plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 1, 0, 75.44245))))
plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 1, 0, 75.44245))))
plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 1, 75.44245))))
plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 1, 75.44245))))
plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 0, 75.44245))))
plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 0, 75.44245))))


#check that it adds up to 1
plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 1, 0, 75.44245)))) + plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 1, 0, 75.44245))),  lower.tail = FALSE) 
plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 1, 0, 75.44245)))) + plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 1, 0, 75.44245))),  lower.tail = FALSE)
plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 1, 75.44245)))) + plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 1, 75.44245))),  lower.tail = FALSE)
plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 1, 75.44245)))) + plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 1, 75.44245))),  lower.tail = FALSE)
plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 0, 75.44245)))) + plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 0, 75.44245))),  lower.tail = FALSE)
plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 0, 75.44245)))) + plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 0, 75.44245))),  lower.tail = FALSE)
```

```{r}
kable(head(newdata5)) %>%
  kable_styling(bootstrap_options = "striped")
```


```{r echo=FALSE}
# We observe that White applicants who meet the credit guidelines with DTI and LTV evaluated at the mean of data set have approximately 13.94% point higher probability of loan approval than similar Black applicants.

0.3235738 - 0.1604961

0.9525447 -0.8891615
```

This table represents the predicted probability of some prototypical individuals across different races from Logit model. This table shows that Black and Hispanic applicants are less likely to receive loan approval in comparison to White applicants. 

We observe that White applicants who meet the credit guidelines with DTI and LTV evaluated at the mean of data set have approximately 16.3% point higher probability of loan approval than similar Black applicants.

We observe that White applicants who do not meet the credit guidelines with DTI and LTV evaluated at the mean of data set have approximately 6.3% point higher probability of loan approval than similar Black applicants. 

```{r}
write_csv(newdata3, "logit_OBRAT.csv")

newdata3$RACE <- ifelse(newdata3$BLACK == 1, "non-Hispanic Black", ifelse(newdata3$HISPAN == 1, "Hispanic", "non-Hispanic White"))
newdata3$RACE  <- as.factor(newdata3$RACE)
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")

ggplot(newdata3, aes(x = OBRAT, y = PredictedProb)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("Predicted probabilities (LOANPRC = 75.44245%)") + 
  facet_wrap(GDLIN ~ ., scales="free") +
  ylab("PredictedProb") + ylim(c(0,1))
```

```{r}
newdata2 <- rbind(newdata34, newdata35, newdata36)
newdata3 <- cbind(newdata2, predict(LogitModel, newdata = newdata2, type = "link", se = TRUE)[1])
newdata3$PredictedProb <- plogis(newdata3$fit)

# newdata3 <- within(newdata3, {
#   PredictedProb <- plogis(fit)
#   # LL <- plogis(fit - (1.96 * se.fit))
#   # UL <- plogis(fit + (1.96 * se.fit))
# })
```

```{r}
write_csv(newdata3, "logit_LOANPRC.csv")

newdata3$RACE <- ifelse(newdata3$BLACK == 1, "non-Hispanic Black", ifelse(newdata3$HISPAN == 1, "Hispanic", "non-Hispanic White"))
newdata3$RACE  <- as.factor(newdata3$RACE)
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")

ggplot(newdata3, aes(x = LOANPRC, y = PredictedProb)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("Predicted probabilities (OBRAT = 32.35767)") + 
  facet_wrap(GDLIN ~ ., scales="free") +
  ylab("PredictedProb") + ylim(c(0,1))
```

### Probit Model

$$Probit(approve) = \beta_0 + \beta_1 * GDLIN + \beta_2 * OBRAT + \beta_3 * BLACK + \beta_4 * HISPAN + \beta_5 * LOANPRC $$

```{r}
ProbitModel <- glm(APPROVE ~  GDLIN + OBRAT + BLACK + HISPAN + LOANPRC, data = data, family = binomial(link = "probit"))
summary(ProbitModel)


#Generate Log-Likelihood
logLik(ProbitModel)

stargazer(ProbitModel,
          type = "text",
          digits = 3,
          header = T,
          title = "Probit Regression")
```

For every one unit change in `OBRAT`, the probit odds of loan approval (versus non loan approval) decreases by `r abs(ProbitModel$coefficients["OBRAT"])`.

For every one unit change in `LOANPRC`, the probit odds of loan approval (versus non loan approval) decreases by `r abs(ProbitModel$coefficients["LOANPRC"])`.

The probit odds of loan approval for applicants that meet credit guidelines increases by  `r abs(ProbitModel$coefficients["GDLIN1"])`.

The probit odds of loan approval for Black applicants decreases by  `r abs(ProbitModel$coefficients["BLACK1"])`.

The probit odds of loan approval for Hispanic applicants decreases by  `r abs(ProbitModel$coefficients["HISPAN1"])`.

For example, for a black person whose credit history meets guidline (GDLIN = 1), loan amount price is 100 (LOANPRC = 100) and other obligations as a percent of total income is none (OBRAT = 0), the probit odds of loan approval is `r (round(1- pnorm(-abs((ProbitModel$coefficients["(Intercept)"] + ProbitModel$coefficients["GDLIN1"] + ProbitModel$coefficients["BLACK1"] + 100 * ProbitModel$coefficients["LOANPRC"]))), 4))*100`%

```{r}
newdata2 <- rbind(newdata31, newdata32, newdata33)
newdata3 <- cbind(newdata2, predict(ProbitModel, newdata2, type = "response", se.fit = TRUE)[1])
write_csv(newdata3, "probit_OBRAT.csv")

newdata4 <- rbind(newdata41, newdata42, newdata43)
newdata6 <- cbind(newdata4, predict(ProbitModel, newdata4, type = "link", se.fit = TRUE)[1], predict(ProbitModel, newdata4, type = "response", se.fit = TRUE)[1])
```

```{r eval=FALSE}
sum(ProbitModel$coefficients * matrix(c(1, 0, 32.35767, 1, 0, 75.44245)))
sum(ProbitModel$coefficients * matrix(c(1, 1, 32.35767, 1, 0, 75.44245)))
sum(ProbitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 1, 75.44245)))
sum(ProbitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 1, 75.44245)))
sum(ProbitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 0, 75.44245)))
sum(ProbitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 0, 75.44245)))
```

```{r}
kable(head(newdata6)) %>%
  kable_styling(bootstrap_options = "striped")
```

```{r}
newdata3$RACE <- ifelse(newdata3$BLACK == 1, "non-Hispanic Black", ifelse(newdata3$HISPAN == 1, "Hispanic", "non-Hispanic White"))
newdata3$RACE  <- as.factor(newdata3$RACE)
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")

ggplot(newdata3, aes(x = OBRAT, y = fit)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("Predicted probabilities (LOANPRC = 75.44245%)") + 
  facet_wrap(GDLIN ~ ., scales="free") +
  ylab("PredictedProb") + ylim(c(0,1))
```


```{r}
newdata2 <- rbind(newdata34, newdata35, newdata36)
newdata3 <- cbind(newdata2, predict(ProbitModel, newdata2, type = "response", se.fit = TRUE)[1])
write_csv(newdata3, "probit_LOANPRC.csv")
```

```{r}
newdata3$RACE <- ifelse(newdata3$BLACK == 1, "non-Hispanic Black", ifelse(newdata3$HISPAN == 1, "Hispanic", "non-Hispanic White"))
newdata3$RACE  <- as.factor(newdata3$RACE)
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")

ggplot(newdata3, aes(x = LOANPRC, y = fit)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("Predicted probabilities (OBRAT = 32.35767)") + 
  facet_wrap(GDLIN ~ ., scales="free") +
  ylab("PredictedProb") + ylim(c(0,1))
```

## Probablities Comparison

```{r}
newdata4$LogitProb <- predict(LogitModel, newdata = newdata4, type = "response", se = FALSE)
newdata4$ProbitProb <- predict(ProbitModel, newdata = newdata4, type = "response", se = FALSE)
```

```{r}
kable(newdata4) %>%
  kable_styling(bootstrap_options = "striped")
```