---
title: "Lending Mortage Analysis - Modeling"
author: "Marjorie Blanco, Joe Thompson, Haodi Tu"
output:
  pdf_document: default
  word_document: default
  html_document: default
subtitle: Exploratory Data Analysis
geometry: left=2cm,right=3cm,top=2cm,bottom=2cm
---

# Introduction 

Early analyses of mortgage approval rates in Boston in 1990 indicated that minorities were more than twice as likely as whites to be denied. Questions remain about whether the difference in loan approval is due to racial bias or other factors that were not considered in the initial analysis such as credit worthiness.  

Consumers expect the mortgage lending process to be impartial about an applicant’s race, based solely on mathematical criteria to evaluate lending risk. If this is not the case, it would be a serious example of racial bias contributing to overall social and economic inequality, as home ownership is an important means of building wealth. We would also expect homeowners to remain longer in their neighborhoods and communities, build strong social ties there, and contribute to local initiatives that create social welfare in those areas. Taken as a whole, bias in mortgage lending could have a cascading effect depressing minority communities on multiple fronts, making economic and social mobility exponentially more difficult for them. 

 

Brief explanation of results... 

 

Econometric Model and Estimation Method 

```{r warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(scipen=999)
library(readr)
library(ggplot2)
library(readxl)
library(kableExtra)
library(dplyr)
library(tidyverse)
library(purrr)
library(ggthemes)
library(aod)
library(lmtest)
library(survey)
library(caret)
library(pscl)
library(survey)
library(stargazer)
library(foreign)
```


```{r  warning=FALSE, message=FALSE, echo=FALSE}
data <- read_excel("MLD/MLD Data File.xls")
# data <- data.frame(data)
data <- data %>% filter(MALE != ".") %>% filter(GDLIN != "666") %>% filter(LOANPRC < 1)
data$LOANPRC <- data$LOANPRC * 100
stargazer(data, type = "text", title="Descriptive statistics", digits=2)
stargazer(data %>% filter(BLACK == 0) %>% filter(HISPAN == 0) %>% select(-c(BLACK, HISPAN)), type = "text", title="Descriptive statistics - non-Hispanic White", digits=2)
stargazer(data %>% filter(BLACK == 1) %>% filter(HISPAN == 0) %>% select(-BLACK), type = "text", title="Descriptive statistics - non-Hispanic Black", digits=2)
stargazer(data %>% filter(BLACK == 0) %>% filter(HISPAN == 1) %>% select(-HISPAN), type = "text", title="Descriptive statistics - Hispanic", digits=2)
```

```{r  warning=FALSE, message=FALSE, echo=FALSE}
# data <- read_excel("MLD Data File.xls")
```

```{r  warning=FALSE, message=FALSE, echo=FALSE}
#http://people.umass.edu/biep640w/pdf/R-for-Logistic-Regression.pdf
#https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf
#https://www.r-bloggers.com/evaluating-logistic-regression-models/
```

```{r  warning=FALSE, message=FALSE, echo=FALSE}
data$GDLIN <- as.factor(data$GDLIN)
data$MARRIED <- as.factor(data$MARRIED)
data$MALE <- as.factor(data$MALE)
data$APPROVE <- as.factor(data$APPROVE)
data$BLACK <- as.factor(data$BLACK)
data$HISPAN <- as.factor(data$HISPAN)
```



```{r eval=FALSE}
#Descriptive statistics by Race:
```

```{r warning=FALSE, message=FALSE, echo=FALSE}
### Overall
summary(data)
data$RACE <- ifelse(data$BLACK == 1, "non-Hispanic Black", ifelse(data$HISPAN == 1, "Hispanic", "non-Hispanic White"))
data$RACE <-as.factor(data$RACE)
data %>% select(-c(BLACK, HISPAN)) %>% split(.$RACE) %>% map(summary) 
```

```{r warning=FALSE, message=FALSE, echo=FALSE}
AVG_LOANPRC <- mean(data$LOANPRC)
AVG_OBRAT <- mean(data$OBRAT)
newdata21 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 192)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(1), HISPAN = factor(0), MARRIED = factor(rep(0:1, each = 96)), LOANPRC = AVG_LOANPRC))
newdata22 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 192)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(1), MARRIED = factor(rep(0:1, each = 96)), LOANPRC = AVG_LOANPRC))
newdata23 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 192)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(0), MARRIED = factor(rep(0:1, each = 96)), LOANPRC = AVG_LOANPRC))
newdata24 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 200)), OBRAT = AVG_OBRAT, BLACK = factor(1), HISPAN = factor(0), MARRIED = factor(rep(0:1, each = 100)), LOANPRC = rep(seq(from = 2.105, to = 100, length.out = 100))))
newdata25 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 200)), OBRAT = AVG_OBRAT, BLACK = factor(0), HISPAN = factor(1), MARRIED = factor(rep(0:1, each = 100)), LOANPRC = rep(seq(from = 2.105, to = 100, length.out = 100))))
newdata26 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 200)), OBRAT = AVG_OBRAT, BLACK = factor(0), HISPAN = factor(0), MARRIED = factor(rep(0:1, each = 100)), LOANPRC = rep(seq(from = 2.105, to = 100, length.out = 100))))
newdata31 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 96)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(1), HISPAN = factor(0), LOANPRC = AVG_LOANPRC))
newdata32 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 96)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(1), LOANPRC = AVG_LOANPRC))
newdata33 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 96)), OBRAT = rep(seq(from = 0, to = 95, length.out = 96)), BLACK = factor(0), HISPAN = factor(0), LOANPRC = AVG_LOANPRC))
newdata34 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 100)), OBRAT = AVG_OBRAT, BLACK = factor(1), HISPAN = factor(0), LOANPRC = rep(seq(from = 2.105, to = 100, length.out = 100))))
newdata35 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 100)), OBRAT = AVG_OBRAT, BLACK = factor(0), HISPAN = factor(1), LOANPRC = rep(seq(from = 2.105, to = 100, length.out = 100))))
newdata36 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 100)), OBRAT = AVG_OBRAT, BLACK = factor(0), HISPAN = factor(0), LOANPRC = rep(seq(from = 2.105, to = 100, length.out = 100))))
newdata41 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 1)), OBRAT = AVG_OBRAT, BLACK = factor(1), HISPAN = factor(0), LOANPRC = AVG_LOANPRC))
newdata42 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 1)), OBRAT = AVG_OBRAT, BLACK = factor(0), HISPAN = factor(1), LOANPRC = AVG_LOANPRC))
newdata43 <- with(data, data.frame(GDLIN = factor(rep(0:1, each = 1)), OBRAT = AVG_OBRAT, BLACK = factor(0), HISPAN = factor(0), LOANPRC = AVG_LOANPRC))
```

# Data 

Our data was gathered in 1990 from lending institutions in Boston, Massachusetts. The full data set contained 1,989 observations of Whites, Blacks, and Hispanics. Given the small number of minority mortgage applications in the measurement period, all Black and Hispanic applications were included in the data set. A random sample of White applicants were added to complete the data. In addition to applicant race, the data includes whether or not a loan was approved, if the applicant’s credit history met the loan guidelines, the applicant’s overall debt obligations compared to their income, whether or not the applicant was male or married, and the loan to value ratio of the desired home. 

Our descriptive statistics show that non-Hispanic whites have the lowest median loan to purchase price percentage (80% versus 87% for Blacks and 89% for Hispanics). Whites also had lower median debt obligations as a percentage of their income (32.5% versus 35% for Blacks and 33.5% for Hispanics). It should be noted that in terms of other debt obligations, the averages for all three groups fall under the guidelines (maximum obligations to income of 36%) that were recommended for mortgages being sold on the secondary market.1 

Overall, we see that 94% of Whites met the loan guidelines and 91% were approved, 73% of Blacks met the loan guidelines and 68% were approved, and 86% of Hispanics met the loan guidelines and 77% were approved.  


## Data Discription

The data set includes the following variables: 

+ APPROVE = 1 if mortgage loan was approved, = 0 otherwise
+ GDLIN = 1 if credit history meets guidelines, = 0 otherwise
+ LOANPRC = loan amount/purchase price
+ OBRAT = other obligations as a percent of total income
+ MALE = 1 if male, = 0 otherwise
+ MARRIED = 1 if married, = 0 otherwise
+ BLACK = 1 if black, = 0 otherwise
+ HISPAN = 1 if Hispanic, = 0 otherwise


# Results 

## Logistic Model

$$log(p/1-p) = \beta_0 + \beta_1 * GDLIN + \beta_2 * OBRAT + \beta_3 * BLACK + \beta_4 * HISPAN + \beta_5 * LOANPRC$$


```{r warning=FALSE, message=FALSE, echo=FALSE}
library(nnet)
LogitModel <- glm(APPROVE ~  GDLIN + OBRAT + BLACK + HISPAN  + LOANPRC , data = data, family = "binomial")
data1  <- data %>% mutate(APPROVE = ifelse(APPROVE == 1, "Yes", ifelse(APPROVE == 0, "No", "Unknown")))
data1$APPROVE <- as.factor(data1$APPROVE)
data1$APPROVE = relevel(data1$APPROVE, ref = "Yes")
multi1 = multinom(APPROVE ~  GDLIN + OBRAT + BLACK + HISPAN  + LOANPRC , data = data1)
#Generate Log-Likelihood
logLik(LogitModel)
stargazer(LogitModel,
          type = "text",
          digits = 3,
          header = T,
          title = "Logistic Regression")
OR <- function(x) exp(x)
stargazer(LogitModel, 
          type="text", 
          apply.coef = OR, 
          title = "Odds Ratio")
stargazer(multi1, type="text")
```

```{r}
or <- exp(coef(LogitModel))
```


### CIs using profiled log-likelihood

```{r}
confint(LogitModel)
```

### CIs using standard errors

```{r}
confint.default(LogitModel)
```

### Odds ratios only

```{r}
exp(coef(LogitModel))
```

### Odds ratios and 95% CI

```{r}
exp(cbind(OR = coef(LogitModel), confint(LogitModel)))
```

Controlling for other variables, applicants who meet the credit guidelines have an estimate of `r or[2]` times higher odds of loan approval than the applicants who do not meet the credit guidelines.  We find that after controlling for all other variables, odds of loan approval for White applicants are 2.46 times and 3.24 times greater in comparison to Blacks and Hispanics respectively.


Holding `GDLIN1`, `LOANPRC` and `Race` constant, for every one unit change in `OBRAT`, the log odds of loan approval (versus non loan approval) decreases by `r round(abs(1- exp(LogitModel$coefficients["OBRAT"])),2)*100`%; 95% confidence interval [CI] = 0.95, 0.99.

Holding `GDLIN1`, `OBRAT` and `Race` constant, for every one unit change in `LOANPRC`, the log odds of loan approval (versus non loan approval) decreases by `r round(abs(1- exp(LogitModel$coefficients["LOANPRC"])),2)*100`%; 95% confidence interval [CI] = 0.97, 1.00.

Holding GDLIN1, OBRAT and Race constant, the log odds of loan approval for applicants that meet credit guidelines over the odds of applicants that do not meet credit guidelines is `r round(abs(exp(LogitModel$coefficients["GDLIN1"])),2)`%, 95% confidence interval [CI] = 27.45, 65.50.

```{r}
newdata2 <- rbind(newdata31, newdata32, newdata33)
newdata3 <- cbind(newdata2, predict(LogitModel, newdata = newdata2, type = "link", se = TRUE)[1])
newdata3$PredictedProb <- plogis(newdata3$fit)
  # LL <- plogis(fit - (1.96 * se.fit))
  # UL <- plogis(fit + (1.96 * se.fit))
#})
newdata4 <- rbind(newdata41, newdata42, newdata43)
newdata5 <- cbind(newdata4, predict(LogitModel, newdata = newdata4, type = "link", se = TRUE)[1])
newdata5$PredictedProb <- plogis(newdata5$fit)
# newdata5 <- within(newdata5, {
#   PredictedProb <- plogis(fit)
  # LL <- plogis(fit - (1.96 * se.fit))
  # UL <- plogis(fit + (1.96 * se.fit))
# })
#predict(LogitModel, newdata = newdata5, type = "response", se = TRUE)
```

```{r eval=FALSE}
sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 1, 0, 75.44245)))
sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 1, 0, 75.44245)))
sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 1, 75.44245)))
sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 1, 75.44245)))
sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 0, 75.44245)))
sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 0, 75.44245)))
plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 1, 0, 75.44245))))
plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 1, 0, 75.44245))))
plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 1, 75.44245))))
plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 1, 75.44245))))
plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 0, 75.44245))))
plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 0, 75.44245))))
#check that it adds up to 1
plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 1, 0, 75.44245)))) + plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 1, 0, 75.44245))),  lower.tail = FALSE) 
plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 1, 0, 75.44245)))) + plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 1, 0, 75.44245))),  lower.tail = FALSE)
plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 1, 75.44245)))) + plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 1, 75.44245))),  lower.tail = FALSE)
plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 1, 75.44245)))) + plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 1, 75.44245))),  lower.tail = FALSE)
plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 0, 75.44245)))) + plogis(sum(LogitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 0, 75.44245))),  lower.tail = FALSE)
plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 0, 75.44245)))) + plogis(sum(LogitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 0, 75.44245))),  lower.tail = FALSE)
```

```{r}
kable(head(newdata5)) %>%
  kable_styling(bootstrap_options = "striped")
```

The predicted probability of approval for a Black person whose credit history meets guidelines (`GDLIN` = 1), at average other debt obligations related to total income (`OBRAT` = 32.36%) and at average loan amount relative to price of the property (`LOANPRC` = 75.44%) is `r round(newdata5$PredictedProb[2], 2)*100`%. White person with the same conditions has higher probability to approve, the predicted probability gap is `r round(newdata5$PredictedProb[6]- newdata5$PredictedProb[2], 2)*100` percentage points.
The predicted probability gap between White and Hispan people under the same background (credit history met guidelines, OBRAT and LOANPRC at the same rate) is `r round(newdata5$PredictedProb[6]- newdata5$PredictedProb[4], 2)*100` percentage points.


```{r echo=FALSE}
# We observe that White applicants who meet the credit guidelines with DTI and LTV evaluated at the mean of data set have approximately 13.94% point higher probability of loan approval than similar Black applicants.
0.3235738 - 0.1604961
0.9525447 -0.8891615
```

This table represents the predicted probability of some prototypical individuals across different races from Logit model. This table shows that Black and Hispanic applicants are less likely to receive loan approval in comparison to White applicants. 

We observe that White applicants who meet the credit guidelines with DTI and LTV evaluated at the mean of data set have approximately 16.3% point higher probability of loan approval than similar Black applicants.

We observe that White applicants who do not meet the credit guidelines with DTI and LTV evaluated at the mean of data set have approximately 6.3% point higher probability of loan approval than similar Black applicants. 

```{r}
write_csv(newdata3, "logit_OBRAT.csv")
newdata3$RACE <- ifelse(newdata3$BLACK == 1, "non-Hispanic Black", ifelse(newdata3$HISPAN == 1, "Hispanic", "non-Hispanic White"))
newdata3$RACE  <- as.factor(newdata3$RACE)
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")
ggplot(newdata3, aes(x = OBRAT, y = PredictedProb)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("Predicted probabilities (LOANPRC = 75.44245%)") + 
  facet_wrap(GDLIN ~ ., scales="free") +
  ylab("PredictedProb") + ylim(c(0,1))
```

The figure above illustrates that the comparison of predicted probability for people with diffierent race when loan price is 75.4% using logit model. When guidelines are not met, gap of predicted probability of approval between White and Hispanic decreases as percentage of other obligations increases, similar with Black. Black and Hispanic have similar predicted probability. On the contrary, gap between White and Hispanic (and Black) increase as percentage of other obligations increases.

```{r}
newdata2 <- rbind(newdata34, newdata35, newdata36)
newdata3 <- cbind(newdata2, predict(LogitModel, newdata = newdata2, type = "link", se = TRUE)[1])
newdata3$PredictedProb <- plogis(newdata3$fit)
# newdata3 <- within(newdata3, {
#   PredictedProb <- plogis(fit)
#   # LL <- plogis(fit - (1.96 * se.fit))
#   # UL <- plogis(fit + (1.96 * se.fit))
# })
```

```{r}
write_csv(newdata3, "logit_LOANPRC.csv")
newdata3$RACE <- ifelse(newdata3$BLACK == 1, "non-Hispanic Black", ifelse(newdata3$HISPAN == 1, "Hispanic", "non-Hispanic White"))
newdata3$RACE  <- as.factor(newdata3$RACE)
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")
ggplot(newdata3, aes(x = LOANPRC, y = PredictedProb)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("Predicted probabilities (OBRAT = 32.35767)") + 
  facet_wrap(GDLIN ~ ., scales="free") +
  ylab("PredictedProb") + ylim(c(0,1))
```

The figure above illustrates that the comparison of predicted probability for people with diffierent race when other obligation is 32.4% using logit model. Similar patterns with figure 1, except the prodicted probability gap is larger between white and other race (Black and Hispanic) when guidelines are not met, and the gap is smaller when guidelines are met.

## Probit Model

$$Probit(approve) = \beta_0 + \beta_1 * GDLIN + \beta_2 * OBRAT + \beta_3 * BLACK + \beta_4 * HISPAN + \beta_5 * LOANPRC $$

```{r}
ProbitModel <- glm(APPROVE ~  GDLIN + OBRAT + BLACK + HISPAN + LOANPRC, data = data, family = binomial(link = "probit"))
#summary(ProbitModel)
#Generate Log-Likelihood
logLik(ProbitModel)
stargazer(ProbitModel,
          type = "text",
          digits = 3,
          header = T,
          title = "Probit Regression")
```

For example, for a black person whose credit history meets guidline (GDLIN = 1), loan amount price is 100 (LOANPRC = 100) and other obligations as a percent of total income is none (OBRAT = 0), the predicted probablity of loan approval is `r (round(pnorm(-abs((ProbitModel$coefficients["(Intercept)"] + ProbitModel$coefficients["GDLIN1"] + ProbitModel$coefficients["BLACK1"] + 100 * ProbitModel$coefficients["LOANPRC"]))), 4))*100`%

```{r}
newdata2 <- rbind(newdata31, newdata32, newdata33)
newdata3 <- cbind(newdata2, predict(ProbitModel, newdata2, type = "response", se.fit = TRUE)[1])
write_csv(newdata3, "probit_OBRAT.csv")
newdata4 <- rbind(newdata41, newdata42, newdata43)
newdata6 <- cbind(newdata4, predict(ProbitModel, newdata4, type = "link", se.fit = TRUE)[1], predict(ProbitModel, newdata4, type = "response", se.fit = TRUE)[1])
```

```{r eval=FALSE}
sum(ProbitModel$coefficients * matrix(c(1, 0, 32.35767, 1, 0, 75.44245)))
sum(ProbitModel$coefficients * matrix(c(1, 1, 32.35767, 1, 0, 75.44245)))
sum(ProbitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 1, 75.44245)))
sum(ProbitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 1, 75.44245)))
sum(ProbitModel$coefficients * matrix(c(1, 0, 32.35767, 0, 0, 75.44245)))
sum(ProbitModel$coefficients * matrix(c(1, 1, 32.35767, 0, 0, 75.44245)))
```

```{r}
kable(head(newdata6)) %>%
  kable_styling(bootstrap_options = "striped")
```

The predicted probability of approval for a Black person whose credit history meets guidelines (`GDLIN` = 1), at average other debt obligations related to total income (`OBRAT` = 32.36%) and at average loan amount relative to price of the property (`LOANPRC` = 75.44%) is `r round(newdata6$fit[2], 2)*100`%. White person with the same conditions has higher probability to approve, the predicted probability gap is `r round(newdata6$PredictedProb[6]- newdata6$PredictedProb[2], 2)*100` percentage points.

```{r}
newdata3$RACE <- ifelse(newdata3$BLACK == 1, "non-Hispanic Black", ifelse(newdata3$HISPAN == 1, "Hispanic", "non-Hispanic White"))
newdata3$RACE  <- as.factor(newdata3$RACE)
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")
ggplot(newdata3, aes(x = OBRAT, y = fit)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Other obligations as a percent of total income") +
  ggtitle("Predicted probabilities (LOANPRC = 75.44245%)") + 
  facet_wrap(GDLIN ~ ., scales="free") +
  ylab("PredictedProb") + ylim(c(0,1))
```

The predicted probability gap between White and Hispan people under the same background (credit history met guidelines, OBRAT and LOANPRC at the same rate) is `r round(newdata6$PredictedProb[6]- newdata6$PredictedProb[4], 2)*100` percentage points.

```{r}
newdata2 <- rbind(newdata34, newdata35, newdata36)
newdata3 <- cbind(newdata2, predict(ProbitModel, newdata2, type = "response", se.fit = TRUE)[1])
write_csv(newdata3, "probit_LOANPRC.csv")
```

```{r}
newdata3$RACE <- ifelse(newdata3$BLACK == 1, "non-Hispanic Black", ifelse(newdata3$HISPAN == 1, "Hispanic", "non-Hispanic White"))
newdata3$RACE  <- as.factor(newdata3$RACE)
newdata3$GDLIN <- ifelse(newdata3$GDLIN == 1, "Meets guidelines", "Does not meet guidelines")
ggplot(newdata3, aes(x = LOANPRC, y = fit)) + 
  geom_line(aes(colour = RACE, linetype = RACE), size = 1)  +
  theme(legend.position="bottom") +
  xlab("Loan amount/purchase price %") +
  ggtitle("Predicted probabilities (OBRAT = 32.35767)") + 
  facet_wrap(GDLIN ~ ., scales="free") +
  ylab("PredictedProb") + ylim(c(0,1))
```

## Probablities Comparison

```{r}
newdata4$LogitProb <- predict(LogitModel, newdata = newdata4, type = "response", se = FALSE)
newdata4$ProbitProb <- predict(ProbitModel, newdata = newdata4, type = "response", se = FALSE)
```

```{r}
kable(newdata4) %>%
  kable_styling(bootstrap_options = "striped")
```

Comparing predicted probabilities of approval from two models, the difference for a Black person whose credit history meets guidelines (`GDLIN` = 1), at average other debt obligations related to total income (`OBRAT` = 32.36%) and at average loan amount relative to price of the property (`LOANPRC` = 75.44%) is `r abs(round(newdata4$LogitProb[2] - newdata4$ProbitProb[2], 3))* 100` percentage points.


# Conclusions

---even if the model shows that other variables have a strong(er) correlation with getting a loan, bias can still be present in the setting of credit history guidelines. Do the guidelines accurately assess the risk of the loans? 


