---
title: "Labor Market Analysis"
author: "Marjorie Blanco, Joe Thomson, Haodi Tu"
output:
  html_document: default
  pdf_document: default
  word_document: default
subtitle: Revised Econometric Model
geometry: left=2cm,right=3cm,top=2cm,bottom=2cm
---

```{r, echo=FALSE}
# Course: ECON 5330
# Title: EDA Assignment
# Date: Feb 5, 2019
# Author: Haodi Tu, Majorie Blanco, Joseph Thomson
```

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(scipen=999)
library(readr)
library(ggplot2)
library(mosaic)
library(dplyr)
library(kableExtra)
library(qwraps2)
library(sjPlot)
library(lmtest)
library(sandwich)
library(olsrr)
```


```{r}
fun_mean <- function(x){
  return(data.frame(y=mean(x),label=round(mean(x,na.rm=T), 2)))}

set_theme_1 <- function() {
  set_theme(
    geom.outline.color = "antiquewhite4", 
    geom.outline.size = 1, 
    geom.label.size = 2,
    geom.label.color = "grey50",
    title.color = "red", 
    title.size = 1, 
    axis.angle.x = 90, 
    axis.textsize.x = 0.9,
    axis.textcolor = "blue", 
    base = theme_bw()
  )
}

set_theme_2 <- function() {
  set_theme(
    geom.outline.color = "antiquewhite4", 
    geom.outline.size = 1, 
    geom.label.size = 2,
    geom.label.color = "grey50",
    title.color = "red", 
    title.size = 1, 
    axis.textsize.x = 0.9,
    axis.textcolor = "blue", 
    base = theme_bw()
  )
}
```

## Research Questions

<!-- https://cran.r-project.org/web/packages/qwraps2/vignettes/summary-statistics.html -->

```{r}
#read data set
ss16ppr <- read_csv("csv_ppr/ss16ppr.csv")
```

```{r}
#select columns of interest
ss16ppr <- ss16ppr %>% select(PERNP, SEX, AGEP, RAC1P, RACWHT, RACBLK, HISP, MAR, SCHL, ENG, COW, WKW, WKHP)
```

```{r}
#convert to numeric data type
ss16ppr$WKHP <- as.numeric(ss16ppr$WKHP)
ss16ppr$PERNP <- as.numeric(ss16ppr$PERNP)
ss16ppr$AGEP <- as.numeric(ss16ppr$AGEP)

#Filters
# AGE: Only between 18 and 64 (inclusive) years of age
ss16ppr <- ss16ppr %>% filter(AGEP >= 18, AGEP <= 64)

# SCHL: High school and above
#       Combine: 16/17 for High school and 18/19 for Some college
ss16ppr <- ss16ppr %>% filter(SCHL >= 16) %>% mutate(SCHL = ifelse(SCHL == 17, 16, SCHL)) %>% mutate(SCHL = ifelse(SCHL == 19, 18, SCHL))

# WKW: Only between 50 and 40 hours
# 1 .50 to 52 weeks worked during past 12 months
# 2 .48 to 49 weeks worked during past 12 months
# 3 .40 to 47 weeks worked during past 12 months
ss16ppr <- ss16ppr %>% filter(between(WKW, 1,3))

# Remove NAs and zero in PERNP
ss16ppr <- ss16ppr %>% filter(!is.na(PERNP))
ss16ppr <- ss16ppr %>% filter(PERNP != 0) 

# Remove outlier
#ss16ppr <-ss16ppr %>% filter(PERNP != max(ss16ppr$PERNP))
# Restirct class of worker to exclude those working without pay and unemployed
# Class of worker
# b N/A (less than 16 years old/NILF who last worked more than 5 years ago or never worked)
# 1 Employee of a private for-profit company or business, or of an individual, for wages,
#    salary, or commissions
# 2 Employee of a private not-for-profit, tax-exempt, or charitable organization
# 3 Local government employee (city, county, etc.)
# 4 State government employee
# 5 Federal government employee
# 6 Self-employed in own not incorporated business, professional practice, or farm
# 7 Self-employed in own incorporated business, professional practice or farm
# 8 Working without pay in family business or farm
# 9 Unemployed and last worked 5 years ago or earlier or never worked
ss16ppr <- ss16ppr %>% filter(between(COW, 1, 7))

# WKHP: worked at least 35 hr per week in a usual work week over this period. 
ss16ppr <- ss16ppr %>% filter(WKHP >= 35)
ss16ppr <- ss16ppr %>% filter(PERNP < 159000) %>% filter(PERNP >= 10000)
```

```{r}
# Decode variables of interest
ss16ppr  <- ss16ppr %>% mutate(Female = ifelse(SEX == 2, 1, 0))
ss16ppr  <- ss16ppr %>% mutate(RACHISP = ifelse(HISP == "01", "Yes", "No"))
ss16ppr  <- ss16ppr %>% mutate(RaceBlack = RACBLK)
ss16ppr  <- ss16ppr %>% mutate(RaceOther = ifelse(RACBLK == 1 | RACWHT == 1, 0, 1))
ss16ppr  <- ss16ppr %>% mutate(RACOTHER = ifelse(RACBLK == 1 | RACWHT == 1, "No", "Yes"))
ss16ppr  <- ss16ppr %>% mutate(RACWHT = ifelse(RACWHT == 1, "Yes", "No"))
ss16ppr  <- ss16ppr %>% mutate(RACBLK = ifelse(RACBLK == 1, "Yes", "No"))


ss16ppr <- ss16ppr %>% mutate(MAR1 = ifelse(MAR == 1, "Yes", "No"))

#code to decode RAC1P
ss16ppr <- mutate(ss16ppr, RAC1P = derivedFactor(
  "White" = RAC1P == 1,
  "Black or African American" = RAC1P == 2,
  "American Indian" = RAC1P == 3,
  "Alaska Native" = RAC1P == 4,
  "American Indian and Alaska Native" = RAC1P == 5,
  "Asian" = RAC1P == 6,
  "Native Hawaiian and Other Pacific Islander" = RAC1P == 7,
  "Some Other " = RAC1P == 8,
  "Two or More Races" = RAC1P == 9,
  method ="first",
  .default = "NA"
))
ss16ppr  <- ss16ppr %>% mutate(Divorced = ifelse(MAR == 3 | MAR == 4, 1, 0))
ss16ppr  <- ss16ppr %>% mutate(NeverMarried = ifelse(MAR == 5, 1, 0))


#code to decode MAR2
ss16ppr <- mutate(ss16ppr, MAR2 = derivedFactor(
  "Married" = MAR == 1 | MAR == 2,
  "Divorced" = MAR == 3 | MAR == 4,
  "Never married" = MAR == 5,
  method ="first",
  .default = "NA"
))

#code to decode MAR
ss16ppr <- mutate(ss16ppr, MAR = derivedFactor(
  "Married" = MAR == 1,
  "Widowed" = MAR == 2,
  "Divorced" = MAR == 3,
  "Separated" = MAR == 4,
  "Never married" = MAR == 5,
  method ="first",
  .default = "NA"
))


#Note: 16 .Regular high school diploma
ss16ppr  <- ss16ppr %>% mutate(SomeCollege = ifelse(SCHL == 18, 1, 0))
ss16ppr  <- ss16ppr %>% mutate(Associate = ifelse(SCHL == 20, 1, 0))
ss16ppr  <- ss16ppr %>% mutate(Bachelor = ifelse(SCHL == 21, 1, 0))
ss16ppr  <- ss16ppr %>% mutate(Master = ifelse(SCHL == 22, 1, 0))
ss16ppr  <- ss16ppr %>% mutate(Professional = ifelse(SCHL == 23, 1, 0))
ss16ppr  <- ss16ppr %>% mutate(Doctoral = ifelse(SCHL == 24, 1, 0))
#code to decode SCHL
ss16ppr <- mutate(ss16ppr, SCHL = derivedFactor(
  "High school diploma" = SCHL == "16",
  "Some college" = SCHL == "18",
  "Associate's degree" = SCHL == "20",
  "Bachelor's degree" = SCHL == "21",
  "Master's degree" = SCHL == "22",
  "Professional degree" = SCHL == "23",
  "Doctorate degree" = SCHL == "24",
  method ="first",
  .default = "NA"
))

# 1 .50 to 52 weeks worked during past 12 months
# 2 .48 to 49 weeks worked during past 12 months
# 3 .40 to 47 weeks worked during past 12 months
#code to decode WKW
ss16ppr <- mutate(ss16ppr, WKW = derivedFactor(
  "50 to 52" = WKW == 1,
  "48 to 49" = WKW == 2,
  "40 to 47" = WKW == 3,
  method ="first",
  .default = "NA"
))

#code to decode SEX
ss16ppr <- mutate(ss16ppr, SEX = derivedFactor(
  "Male" = SEX == 1,
  "Female" = SEX == 2,
  method ="first",
  .default = "NA"
))
```


1) How do earnings vary by education level?
2) How does the premium for education vary by gender?

## Revised Model

$Earning = \beta_0 + Divorced * \beta_1 + NeverMarried * \beta_2 + Female * \beta_3 + RaceBlack * \beta_4 + RaceOther * \beta_5 +$ 

$SomeCollege * \beta_6 + Associate * \beta_7 + Bachelor * \beta_8 + Master * \beta_9 + Professional * \beta_10 + Doctoral * \beta_11 + Age * \beta_12$

```{r}
ss16ppr  <- ss16ppr %>% mutate(Widowed = ifelse(MAR == "Widowed", 1, 0))
ss16ppr  <- ss16ppr %>% mutate(Divorced = ifelse(MAR == "Divorced", 1, 0))
ss16ppr  <- ss16ppr %>% mutate(Separated = ifelse(MAR == "Separated", 1, 0))
ss16ppr  <- ss16ppr %>% mutate(NeverMarried = ifelse(MAR == "Never married", 1, 0))


Model_1 <- lm(PERNP ~ Widowed + Divorced + Separated + NeverMarried + Female + RaceBlack + RaceOther + SomeCollege + Associate + Bachelor + Master + Professional + Doctoral + AGEP, data = ss16ppr)
summary(Model_1)
```

### Interpretation

+ Coefficients Explanation 
+ Holding gender, race, education and age constant, married people makes $`r abs(round(Model_1$coefficients["Widowed"], 2))` less than people who widowed on average.
+ Holding gender, race, education and age constant, married people makes $`r abs(round(Model_1$coefficients["Divorced"], 2))` more than people who divorced on average.
+ Holding gender, race, education and age constant, married people makes $`r abs(round(Model_1$coefficients["Separated"], 2))` more than people who separated on average.
+ Holding gender, race, education and age constant, married people makes $`r abs(round(Model_1$coefficients["NeverMarried"], 2))` more than people who never married on average.
+ Holding marriage, race, education and age constant, male makes $`r abs(round(Model_1$coefficients["Female"], 2))` more than female on average.
+ Holding marriage, gender, education and age constant, White makes $`r abs(round(Model_1$coefficients["RaceBlack"], 2))` more than Black on average.
+ Holding marriage, gender, education and age constant, White makes $`r abs(round(Model_1$coefficients["RaceOther"], 2))` more than Other race on average.
+ Holding marriage, gender, race and age constant, people have high school education makes $`r abs(round(Model_1$coefficients["SomeCollege"], 2))` less than people have some college education on average.
+ Holding marriage, gender, race and age constant, people have high school education makes $`r abs(round(Model_1$coefficients["Associate"], 2))` less than people have associate education on average.
+ Holding marriage, gender, race and age constant, people have high school education makes $`r abs(round(Model_1$coefficients["Bachelor"], 2))` less than people have bachelor’s degree on average.
+ Holding marriage, gender, race and age constant, people have high school education makes $`r abs(round(Model_1$coefficients["Master"], 2))` less than people have master’s degree on average.
+ Holding marriage, gender, race and age constant, people have high school education makes $`r abs(round(Model_1$coefficients["Professional"], 2))` less than people have Professional education on average.
+ Holding marriage, gender, race and age constant, people have high school education makes $`r abs(round(Model_1$coefficients["Doctoral"], 2))` less than people have doctor’s degree on average.
+ Holding marriage, gender, race and education constant, people make $`r abs(round(Model_1$coefficients["AGEP"], 2))` more as age increases on average between the age of 18 to 64.


### Model Diagnostics

```{r}
plot_model(Model_1, group.terms = c(1,1,1,1,2,3,3,4,4,4,4,4,4,5))

p <- plot_model(Model_1, type = "diag")
```

```{r echo=FALSE}
kable(ols_vif_tol(Model_1))

set_theme_1()
p[1]
```

A VIF chart shows that there is no correlation among the kth predictor and the remaining predictor variables.

```{r echo=FALSE}
set_theme_2()
p[2]
```

The plot shows that the distribution of the residuals are approximately normally distributed.  The “thin tails” correspond to the first quantiles occurring at less than expected values and the last quantiles occurring at larger than expected values. 

```{r echo=FALSE}
set_theme_1()
p[3]
```

The plot shows that the distribution of the residuals are approximately normally distributed.

```{r echo=FALSE}
set_theme_1()
p[4]
```

The plot shows heteroscedasticity as the variance of the residuals are not constant. 

### Breusch-Pagan test

$H0$: Variance is unchanging in the residual
$H1$: Variance is changing in the residual

```{r}
#Conduct BPG Test
bptest(Model_1)
```

The p-Value < 0.05 which indicates that the null hypothesis (the variance is unchanging in the residual) can be rejected and therefore heterscedasticity exists.

### Revised Model wit robust standard errors

```{r}
bptestequation = lm(residuals(Model_1)*residuals(Model_1) ~ Widowed + Divorced + Separated +
                      NeverMarried + Female + RaceBlack + RaceOther + SomeCollege + Associate + 
                      Bachelor + Master + Professional + Doctoral + AGEP, data=ss16ppr)
summary(bptestequation) #note: BP = n*RSquared of model with squared residuals as dependent variable 

#Generate the Variance Covariance Matrix of the Parameter Estimates
vcovHC(Model_1, type = "HC") #the diagonal elements are the variances of the parameter estimates

#Generate the Robust standard errors and print them on screen 
sandwich_se <- diag(vcovHC(Model_1, type = "HC"))^0.5
sandwich_se
```

## Revised Logarithmic Model

```{r}
#Estimate Logarithmic Model with Age in Quadratic Form
LogModel_1 = lm(log(PERNP, base = exp(1)) ~ Widowed + Divorced + Separated +
                  NeverMarried + Female + RaceBlack + RaceOther + SomeCollege + Associate + 
                  Bachelor + Master + Professional + Doctoral + AGEP, data=ss16ppr)

summary(LogModel_1)
```

### Interpretation

+ Coefficients Explanation 

+ Holding gender, race, education and age constant, married people makes $`r abs(round((exp(LogModel_1$coefficients["Widowed"]) -1) * 100, 2))`% less than people who widowed on average.
+ Holding gender, race, education and age constant, married people makes $`r abs(round((exp(LogModel_1$coefficients["Divorced"]) -1) * 100, 2))`% more than people who divorced on average.
+ Holding gender, race, education and age constant, married people makes $`r abs(round((exp(LogModel_1$coefficients["Separated"]) -1) * 100, 2))`% more than people who separated on average.
+ Holding gender, race, education and age constant, married people makes $`r abs(round((exp(LogModel_1$coefficients["NeverMarried"]) -1) * 100, 2))`% more than people who never married on average.
+ Holding marriage, race, education and age constant, male makes $`r abs(round((exp(LogModel_1$coefficients["Female"]) -1) * 100, 2))`% more than female on average.
+ Holding marriage, gender, education and age constant, White makes $`r abs(round((exp(LogModel_1$coefficients["RaceBlack"]) -1) * 100, 2))`% more than Black on average.
+ Holding marriage, gender, education and age constant, White makes $`r abs(round((exp(LogModel_1$coefficients["RaceOther"]) -1) * 100, 2))`% more than Other race on average.
+ Holding marriage, gender, race and age constant, people have high school education makes $`r abs(round((exp(LogModel_1$coefficients["SomeCollege"]) -1) * 100, 2))`% less than people have some college education on average.
+ Holding marriage, gender, race and age constant, people have high school education makes $`r abs(round((exp(LogModel_1$coefficients["Associate"]) -1) * 100, 2))`% less than people have associate education on average.
+ Holding marriage, gender, race and age constant, people have high school education makes $`r abs(round((exp(LogModel_1$coefficients["Bachelor"]) -1) * 100, 2))`% less than people have bachelor’s degree on average.
+ Holding marriage, gender, race and age constant, people have high school education makes $`r abs(round((exp(LogModel_1$coefficients["Master"]) -1) * 100, 2))`% less than people have master’s degree on average.
+ Holding marriage, gender, race and age constant, people have high school education makes $`r abs(round((exp(LogModel_1$coefficients["Professional"]) -1) * 100, 2))`% less than people have Professional education on average.
+ Holding marriage, gender, race and age constant, people have high school education makes $`r abs(round((exp(LogModel_1$coefficients["Doctoral"]) -1) * 100, 2))`% less than people have doctor’s degree on average.
+ Holding marriage, gender, race and education constant, people make $`r abs(round((exp(LogModel_1$coefficients["AGEP"]) -1) * 100, 2))`% more as age increases on average between the age of 18 to 64.

### Model Diagnostics

```{r}
plot_model(LogModel_1, group.terms = c(1,1,1,1,2,3,3,4,4,4,4,4,4,5))

p <- plot_model(LogModel_1, type = "diag")
```

```{r echo=FALSE}
kable(ols_vif_tol(LogModel_1))

set_theme_1()
p[1]
```

A VIF chart shows that there is no correlation among the kth predictor and the remaining predictor variables.

```{r echo=FALSE}
set_theme_2()
p[2]
```

The plot shows that the distribution of the residuals are approximately normally distributed.  The “thin tails” correspond to the first quantiles occurring at less than expected values and the last quantiles occurring at larger than expected values. 

```{r echo=FALSE}
set_theme_1()
p[3]
```

The plot shows that the distribution of the residuals are approximately normally distributed.

```{r echo=FALSE}
set_theme_1()
p[4]
```

The plot shows heteroscedasticity as the variance of the residuals are not constant. 

```{r echo=FALSE, include=FALSE}
library(stargazer)

# gather robust standard errors in a list
rob_se <- list(sqrt(diag(vcovHC(Model_1, type = "HC1"))),
               sqrt(diag(vcovHC(LogModel_1, type = "HC1"))))

# generate a LaTeX table using stargazer
stargazer(Model_1, LogModel_1,
          type = "latex",
          se = rob_se,
          digits = 3,
          header = F,
          column.labels = c("(I)", "(II)"),  
          out="RevisedModels.htm")
```
